### 2023-11-01 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@COMPAS-Surrogate}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2023-11-01 src/tex/ms.tex
Update ms.tex

@@
 
+\section{Introduction} \label{sec:intro}
+This work builds of Riley et al 2022
+
+Astrophysical processes that govern stellar, binary system evolution, and the production of binary black holes (BBH) that produce gravitational waves (GW) detectable by current instrumentation (e.g.
+aLIGO, VIRGO, KAGRA, the \lvk) are very uncertain.
+Over the next few years, we will accumulate a population of several hundred BBH mergers. 
+We can use the population of mergers to constrain some parameters describing astrophysical processes, such as the star formation rate in the early universe. 
+In this work, we use \compas, a rapid stellar population synthesis tool, to model the universe with different star formation rates (SFR) and try to place constraints on the SFR that can reproduce the \lvk\ detections.
+
+The remainder of the paper is organised as follows.
+Section~\ref{sec:method} outlines...
+Results are summarised in Section~\ref{sec:results}.
+The data products, and software to reproduce the results available online as supplementary materials (\projectUrl).
+Finally, we discuss caveats and provide concluding remarks in Section~\ref{sec:conclusion}.
+
+\section{Method} \label{sec:method}
+
+The \lvk\ has detected $\sim90$ GW signals from a population of merging BBHs. 
+Each BBH has some parameters $\vec{\theta}$ that describe the system, such as the BBH's chirp-mass $\mathcal{M}$, spin magnitude $\chi$, spin angle $\theta$, and redshift $z$.
+On the other hand, \compas\ can generate populations of merging BBHs using different initial star formation parameters $\vec{\lambda}$.
+
+Therefore, we can use \compas\ to determine the posterior on $\vec{\lambda}$ given \lvk's $N_{\rm obs}$ observed BBHs and their parameters $\vec{D} = \{ p_i(\vec{\theta}|h), ...\}$ (where $h$ is the strain data):
+\begin{equation}
+    p(\vec{\lambda}| \vec{D}) \sim \pi(\vec{\lambda})\ \mathcal{L}(\vec{D}| \vec{\lambda})\ .
+\end{equation}
+
+Here, $\mathcal{L}(\vec{D}| \vec{\lambda})$ can be written as:
+\begin{equation}
+    \ln \mathcal{L}(\vec{D}| \vec{\lambda}) = \ln \mathcal{L}(N_{\rm obs}|\vec{\lambda}) + \sum^{N_{\rm obs}}_{i=1} \ln \mathcal{L}(D_i|\vec{\lambda}) \ .
+\end{equation}
+
+Expected gains
+
+\section{Toy Model}
+
+In this section, we delve into a toy-model problem of just using the \lvk\ chirp-masses and \compas\ chirp-masses, marginalising over the redshift.
+
+In this case, we create a surrogate for $\mathcal{L}$.
+
+\section{Results}\label{sec:results}
+Details on the results
+
+\section{Discussion and caveats}\label{sec:conclusion}
+Some concluding discussions
+
+\section{Data and software availability}\label{sec:data}
+Code data can be found at \projectGit
+
+%%%%%%%%%%%%%%%%%%%%
+
+\section*{Acknowledgments}{
+
+We gratefully acknowledge the Swinburne Supercomputing OzSTAR Facility for computational resources.
+All analyses (including test and failed analyses) performed for this study used $XX$K core hours on OzSTAR.
+This would have amounted to a carbon footprint of ${\sim XX{\text{t CO}_2}}$~\citep{greenhouse, energy_to_co2_converter}.
+Thankfully, as OzSTAR is powered by wind energy from Iberdrola Australia; the electricity for computations produces negligible carbon waste.
+
+A.V. is supported by the Australian Research Council (ARC) Centre of Excellence CE170100004.
+
+}
+
+\vspace{5mm}
+\facilities{LIGO-VIRGO-KAGRA}
+
+\software{
+\python~\citep{pythonForScientificComputing,pythonForScientists},
+\astropy~\citep{astropy},
+\arviz~\citep{arviz_2019},
+\exoplanet~\citep{Foreman-Mackey:2021:JOSS},
+\lightkurve~\citep{LightkurveCollaboration:2018:ascl},
+\starry~\citep{Luger:2019:AJ},
+\celerite~\citep{Foreman-Mackey:2017:ascl},
+\pymc~\citep{Salvatier:2016:ascl},
+\numpy~\citep{numpy},
+\scipy~\citep{SciPy},
+\pandas~\citep{pandas},
+\matplotlib~\citep{matplotlib},
+\corner~\citep{corner},
+\sphinx~\citep{sphinx_doc},
+\jupyter~\citep{jupyter},
+\jupyterbook~\citep{jupyter_book}.
+}
+
 \bibliography{bib}

### 2024-06-25 src/tex/ms.tex
add more plots

--- 2023-11-01 src/tex/ms.tex
@@
 
+\usepackage{natbib}
+\usepackage[utf8]{inputenc}
+\usepackage{amsmath}
+\usepackage{amsmath,units}
+\usepackage{amsmath,bm}
+\usepackage{booktabs}
+\usepackage{makecell}
+\usepackage{xcolor}
+\usepackage{graphicx}
+\usepackage{acronym}
+\usepackage{xspace}
+\usepackage{enumitem}
+\usepackage{calc}
+\usepackage[caption=false]{subfig}
+\usepackage{import}
+\usepackage{mathtools}
+\usepackage{showyourwork}
+
+\setlength{\marginparwidth}{1.5cm}
+
+% table column types
+\usepackage{array}
+\newcolumntype{E}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{F}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{G}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+
+% abbreviations/acronyms
+\acrodef{ANN}{artificial neural network}
+\acrodef{DCO}{double compact object}
+\acrodef{MCMC}{Markov Chain Monte Carlo}
+\acrodef{MSSFR}{metallicity-specific star formation rate}
+\acrodef{LVK}{LIGO--VIRGO--KAGRA}
+
+\input{custom_commands}
+
 % Begin!
@@
 
-% Title
-\title{An open source scientific article}
+\title{\resp{COMPAS Star-formation LnL GP surrogate}}
 
-% Author list
-\author{@COMPAS-Surrogate}
+% Authors
+\author{Avi Vajpeyi}
+\affiliation{UoA}
+\affiliation{NZ Gravity}
+\affiliation{ARC Centre of Excellence for Gravitational Wave Discovery -- OzGrav, Australia}
 
-% Abstract with filler text
+\author{Jeff Riley}
+\affiliation{School of Physics and Astronomy, Monash University, Clayton, Victoria 3800, Australia}
+\affiliation{ARC Centre of Excellence for Gravitational Wave Discovery -- OzGrav, Australia}
+\author{Ilya Mandel}
+\affiliation{School of Physics and Astronomy, Monash University, Clayton, Victoria 3800, Australia}
+\affiliation{ARC Centre of Excellence for Gravitational Wave Discovery -- OzGrav, Australia}
+\author{Chayan Chatterjee}
+\affiliation{}
+
+\correspondingauthor{Avi Vajpeyi}
+\email{avi.vajpeyi@auckland.ac.nz}
+
+%%%%
+
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
+    We're using GPs to build surrogate likelihoods for COMPAS star formation parameters, given the LVK observed dataset.
 \end{abstract}
 
-% Main body with filler text
+\keywords{surrogate model -- inference -- binaries -- stars: evolution -- gravitational waves -- machine learning -- Gaussian Processes}
+
 \section{Introduction}
 \label{sec:intro}
+Over the past decade, the improvements in detector sensitivity have led to an exponential increase in the rate of GW detections.
+With next-generation detectors on the horizon (such as CE, ET, LISA), the number of GW detections will increase even more.
+The most recent catalogue of GW contains
+XX events, the majority of them being Binary Black Holes
+(BBHs), with XX public alerts for GW that have yet to be confirmed as astrophysiocal events.
+As more GW events are observed, more features
+in the distribution of GW events await to be discovered.
+One of the most common ways to understand the observed GW population is through phenomenological modelling.
+However, as the size of the catalogues increases, so does the catalogue complexity, making it challenging to develop interpretable, parameterisable phenomenological models, flexible enough to explain the features in the catalogue.
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
+Alternatively, we can consider using forward models to simulate catalogues of GW events and compare the simulations to observed catalogues.
+The forward modelling approach would provide astronomers with constraints on easy-to-interpret physical parameters that describe the initial conditions of the population of merging binaries, such as distributions describing the initial stellar mass, metallicity, and binary separation.
+These parameters also govern simulated physical processes like wind mass loss, supernova characteristics, and binary interactions such as mass transfer and common envelope efficiency.
+However, the generation of synthetic catalogues from the various rapid population sysnthesis software packages can be expensive, making brute force methods that explore all possible parameter values expensive (for example, see  Broekgaarden et al.
+(2022)).
+Another alternative, as proposed by Riley et al. (2023) involves building surrogate tools to bridge the computational complexities of simulating a state space large enough
+to allow us to infer astrophysical constraints from the observed catalogs.
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+Riley et al. (2023) generated a small number of ``training" states using a
+targeted selection of initial conditions and evolutionary
+parameters.
+The training states were then used to teach an interpolant how to
+map initial conditions and evolutionary parameters to
+the final state, thus avoiding the computational time
+of calculating the intermediate steps.
+
+In their work, they explored four parameters governing the cosmic metallicity-specific star formation rate (MSSFR) in the model of Neijssel et al. (2019).
+
+We build upon Riley at al. (2023)'s work, and
+
+The remainder of this paper is organised as follows.
+Section 2 presents a description of the tool we construct
+for this proof-of-concept study, and the method used
+to train the tool.
+We present and discuss our results in Section 3.
+We provide some concluding remarks in  Section 4
+
+%--------------------------------------------------------
+\section{Method}
+\subsection{Gaussian process and Bayesian Optimisation}
+Talk about bayesian optimisation
+
+\subsection{MSSFR parameters}\label{sec:method_parameters}
+
+The parameters we chose to vary for this study are four free parameters for the calculation of the \ac{MSSFR} in the phenomenological model of \citet{Neijssel_2019}.
+In this model, the \ac{MSSFR} is split into two parts, the star formation rate (SFR) and the metallicity distribution:
+
+\begin{equation}
+  \frac{\diff^3 \MSFR}{\diff \ts \diff \Vc \diff Z}(z) =
+  \frac{\diff^2 \MSFR}{\diff \ts \diff \Vc }(z)
+  \times
+  \frac{\diff P }{\diff Z}(z),
+\end{equation}
+
+\noindent
+where the SFR is given by:
+
+\begin{equation}
+  \frac{\diff^2 \MSFR}{\diff \ts \diff \Vc} =
+  a\frac{(1 + z)^b}{1 + (\frac{(1 + z)}{c})^d}\,
+  \mathrm{M}_\odot\ \mathrm{yr}^{-1}\ \mathrm{Mpc}^{-3},
+\end{equation}
+
+\bigskip\noindent
+and we use a log-normal distribution in metallicity at each redshift (cf.~the skewed log-normal model proposed by \citet{vanSon_2022}):
+
+\begin{equation}
+  \frac{\diff P }{\diff Z}(z) =
+  \frac{1}{Z\sigma\sqrt{2\pi}}\, \exp\bigg[{-\frac{(\ln Z-\ln \langle Z \rangle +\sigma^2/2)^2}{2\sigma^2}}\bigg],
+\end{equation}
+
+\bigskip\noindent
+with a redshift-independent standard deviation $\sigma$ in $\ln Z$ space around a redshift-dependent mean $\mu$ of $\ln Z$ given by:
+
+\begin{equation}
+  \langle Z \rangle =
+  \exp[\mu + \sigma^2/2],
+\end{equation}
+
+\bigskip\noindent
+with mean metallicity parametrised as in \citet{Langer_2006}:
+
+\begin{equation}
+  \langle Z(z) \rangle =
+  Z_{0}10^{\alpha{z}}
+\end{equation}
+
+\bigskip\noindent
+We vary SFR parameters $a$ and $d$, and the metallicity distribution parameters $\alpha$ and $\sigma$, in this study, while fixing $b=2.77$, $c=2.9$, and $Z_0=0.035$.
+\resp{\citet{Neijssel_2019} demonstrated that a range of star-formation history models can be mimicked by varying only this subset of \ac{MSSFR} parameters.
+We therefore choose to vary the same parameters for consistency with their analysis.
+Moreover, reducing the number of parameters allows us to limit computational complexity for this proof-of-principle study.}
+
+\subsection{Cosmic integration}
+Talk about how we run cosmic int,
+show plots of Mcz grid for different SF
+
+\subsection{Bayesian Model}
+We define the likelihood $\mathcal{L}$ that we will see the data $\mathcal{D}$ for a particular set of parameters $\lambda$ as
+
+\begin{equation}
+  \log{\mathcal{L}\left( \mathcal{D} |\lambda\right)} = \log{\Lp\left(N_{\mathrm{obs}}|\lambda\right)} +
+  \sum_{i=1}^{N_\mathrm{obs}} \log{\Le(D_i|\lambda)},
+  \label{eq:methods_likelihood}
+\end{equation}
+
+\bigskip\noindent
+where $N_{\mathrm{obs}}$ is the number of observations in the data $\mathcal{D} \equiv \{D_1, D_2, ... D_{N_{\mathrm{obs}}}\}$.
+
+The first term of Equation~\ref{eq:methods_likelihood} is a Poisson likelihood on there being $N_\mathrm{obs}$ detections,
+
+\begin{equation}
+  %\notag
+  \log{ \Lp\left(N_{\mathrm{obs}}|\lambda\right)} = - \mu(\lambda) + N_{\mathrm{obs}}\log\mu(\lambda) ,
+  \label{eq:methods_likelihood_term1}
+\end{equation}
+
+\bigskip\noindent
+where $\mu(\lambda)$ is the expected number of observations and we omit terms that depend on the data only and therefore disappear on normalisation, such as $\log(N_{\mathrm{obs}}!)$ and permutation coefficients.
+
+The second term of Equation~\ref{eq:methods_likelihood} is comprised of a product of the probabilities of individual event detections,
+
+\begin{equation}
+  %\notag
+  \Le(D_i|\lambda) = \pg(z=z_i,\mathcal{M}=\mathcal{M}_{i}|\lambda),
+  \label{eq:methods_likelihood_term2}
+\end{equation}
+
+\bigskip\noindent
+where $\pg(z=z_i,\mathcal{M}=\mathcal{M}_{i}|\lambda)$ is the entry in the probability distribution matrix $\pg(z,\mathcal{M}|\lambda)$ in the $z$ and $\mathcal{M}$ bin of the observed $i^{th}$ event.
+The probability distribution matrix $\pg(z,\mathcal{M}|\lambda)$ is constructed by dividing the entries in the matrix of expected detections for a given $\lambda$, generated by multiplying the surrogate model detection rate matrix by the run duration, by the sum of the entries in the matrix, $\mu$.
+
+\subsection{Using gravitational wave observations}\label{sec:method_LVK}
+
+The catalogue of LIGO, Virgo, and KAGRA (LVK) gravitational wave observations (\citet{GWTC-2_1_zenodo}, \citet{GWTC-3_zenodo}) provides us with a number of merger detections from a population of \acp{DCO}.
+For each LVK event we know, among other things, the chirp mass of the merging binary and the merger redshift, up to a measurement uncertainty.
+We only include confident detections with a minimum astrophysical probability $p_{astro} \ge 0.95$; although less confident events could be included \citep[e.g.][]{Farr_2015}, they will typically contribute little information due to greater measurement uncertainties, so we omit them for this proof-of-concept study.
+
+Once we validate our method, we can use our surrogate model and the LVK data to infer real physical constraints on the parameters under study, with the important caveat that we have fixed the astrophysical parameters, and any errors in those could bias the inferred \ac{MSSFR} model values.
+
+\subsubsection{LVK likelihood calculation}\label{sec:method_LVK_likelihood_calculation}
+
+ In practice, for LVK events, we may not know the values $z_i, \Mc_{i}$ and hence the probability $\Le(z=z_i,\Mc=\Mc_{i}|\lambda)$ for the observed $i^{th}$ event in Equation~\ref{eq:methods_likelihood_term2} perfectly, but may only have $K$ samples from a posterior $p(z_i,\Mc_{i}|D_i)$.
+For LVK data, we substitute the following for Equation~\ref{eq:methods_likelihood_term2}:
+
+\begin{equation}
+  \Le(D_i|\lambda) = \frac{1}{K} \sum_{k=1}^K \frac{\pg(z=z_i^k, \Mc=\Mc_{i}^k|\lambda)}{\pi(z_i^k,\Mc_i^k)},
+  \label{eq:methods_likelihood_term2_LIGO}
+\end{equation}
+
+\bigskip\noindent
+where the subscript $k$ refers to the $k^{th}$ posterior sample among $K$ available samples for event $i$ and $\pi(z,\Mc)$ is the prior used by LVK data analysts in the original inference on the observed events \citep[see, e.g.][]{Mandel_2019}.
+
+In practice, if the number of bins where an event has posterior support is much smaller than the number of samples $K$, it may be more efficient to pre-compute the weight that each observed event $i$ contributes to each of the bins $b$, $w_{i,b}$, which is independent of $\lambda$:
+
+\begin{equation}
+    w_{i,b} = \frac{1}{K} \sum^K_{k=1} \frac{I(\{z^k_i,\Mc^k_i\} \in b)}{\pi(z_i^k,\Mc_i^k)} \ ,
+\end{equation}
+where $I(\{z^k_i,\Mc^k_i\} \in b$ is an indicator function that evaluates to 1 if sample $\{z^k_i,\Mc^k_i\}$ falls into bin $b$ and $0$ otherwise.
+Then the probability $\Le(D_i|\lambda)$ can be evaluated as a sum over bins,
+
+\begin{equation}
+    \Le(D_i|\lambda) = \sum_b \pg(z^b, \Mc^b | \lambda) w_{i,b}
+\end{equation}
+
+\subsection{Diagnostics}
+variable LnL and Kl divergence
+
+%--------------------------------------------------------
+\section{Toy model}
+Include corners of variable LnL + normal LnL
+Include KL divergence (low number points --> high points)
+Include PP plot
+
+\begin{figure}[ht!]
+    \script{plot_kl_distances.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/kl_distances.pdf}
+        \caption{
+            KL-Distances of the posteriors obtained with XX, compared against to those from YY.
+        }
+        \label{fig:kl_distances}
+    \end{centering}
+\end{figure}
+
+\begin{figure}[ht!]
+    \script{plot_pp_plot.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/pp_plot.pdf}
+        \caption{
+            PP-plots from simulation study.
+        }
+        \label{fig:pp_plot}
+    \end{centering}
+\end{figure}
+
+%--------------------------------------------------------
+\section{LVK analysis}
+\subsection{Data}
+
+\begin{figure}[ht!]
+    \script{plot_ogc4_events.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/ogc4_events.pdf}
+        \caption{
+            The events and uncertainty as predicted by Ogc4
+        }
+        \label{fig:ogc4_events}
+    \end{centering}
+\end{figure}
+
+\begin{figure}[ht!]
+    \script{plot_ogc4_weights.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/ogc4_weights.pdf}
+        \caption{
+            Weights generated using the OGC4 posteriors
+        }
+        \label{fig:ogc4_weights}
+    \end{centering}
+\end{figure}
+
+Describe OGC4 data
+
+\subsection{Results}
+Describe LVK results, compare with Riley et al.
+
+%--------------------------------------------------------
+\section{Discussion}
+
+%%% DATA AVAILABILITY %%%
+\section*{Data availability}
+The data underlying this article is available in the Zenodo deposit for this work. LINK TO ZENODO.
+
+\section*{Acknowledgements}
+We gratefully acknowledge the Swinburne Supercomputing OzSTAR Facility for computational resources.
+All analyses (including test and failed analyses) performed for this study used $XX$K core hours on OzSTAR.
+This would have amounted to a carbon footprint of ${\sim XX{\text{t CO}_2}}$~\citep{greenhouse, energy_to_co2_converter}.
+Thankfully, as OzSTAR is powered by wind energy from Iberdrola Australia; the electricity for computations produces negligible carbon waste.
+A.V. is supported by the Australian Research Council (ARC) Centre of Excellence CE170100004.
+
+4-OGC.
+IM is a recipient of the Australian Research Council Future Fellowship FT190100574.
+AV is XX.
+Discussions with Renate, Matt, Kate.
+
+\vspace{5mm}
+\facilities{LIGO-VIRGO-KAGRA}
+
+\software{
+\python~\citep{pythonForScientificComputing,pythonForScientists},
+\astropy~\citep{astropy},
+\arviz~\citep{arviz_2019},
+\exoplanet~\citep{Foreman-Mackey:2021:JOSS},
+\lightkurve~\citep{LightkurveCollaboration:2018:ascl},
+\starry~\citep{Luger:2019:AJ},
+\celerite~\citep{Foreman-Mackey:2017:ascl},
+\pymc~\citep{Salvatier:2016:ascl},
+\numpy~\citep{numpy},
+\scipy~\citep{SciPy},
+\pandas~\citep{pandas},
+\matplotlib~\citep{matplotlib},
+\corner~\citep{corner},
+\sphinx~\citep{sphinx_doc},
+\jupyter~\citep{jupyter},
+\jupyterbook~\citep{jupyter_book}.
+}
 
@@
 
+\appendix
+
+\section{$\pi(\mathcal{M}, D_{C})$ to $\pi(\mathcal{M}, z)$}
+
+See \cite{Hogg:1999:arXiv} for transforms
+
+\begin{equation}
+\Omega_{\rm M}\equiv\frac{8\pi\,G\,\rho_0}{3\,H_0^2}
+\end{equation}
+\begin{equation}
+\Omega_{\Lambda}\equiv\frac{\Lambda\,c^2}{3\,H_0^2}
+\end{equation}
+\begin{equation}
+    D_H = c / H_0
+\end{equation}
+
+\begin{equation}
+\label{eq:ez}
+E(z)\equiv\sqrt{\Omega_{\rm M}\,(1+z)^3+\Omega_k\,(1+z)^2+\Omega_{\Lambda}}
+\end{equation}
+
+\begin{equation}
+D_{\rm C} = D_{\rm H}\,\int_0^z\frac{dz'}{E(z')}
+\end{equation}
+
+The comoving distance between two events at the same redshift or
+distance but separated on the sky by some angle $\delta\theta$ is
+$D_{\rm M}\,\delta\theta$ and the transverse comoving distance $D_{\rm
+M}$ (so-denoted for a reason explained below) is simply related to the
+line-of-sight comoving distance $D_{\rm C}$:
+\begin{equation}
+D_{\rm M} = \left\{
+\begin{array}{ll}
+D_{\rm H}\,\frac{1}{\sqrt{\Omega_k}}\,\sinh\left[\sqrt{\Omega_k}\,D_{\rm C}/D_{\rm H}\right] & {\rm for}~\Omega_k>0 \\
+D_{\rm C} & {\rm for}~\Omega_k=0 \\
+D_{\rm H}\,\frac{1}{\sqrt{|\Omega_k|}}\,\sin\left[\sqrt{|\Omega_k|}\,D_{\rm C}/D_{\rm H}\right] & {\rm for}~\Omega_k<0
+\end{array}
+\right.
+\end{equation}
+
+Angular diameter distance is related to the
+transverse comoving distance by
+\begin{equation}
+D_{\rm A} = \frac{D_{\rm M}}{1+z}
+\end{equation}
+
+The comoving volume element in solid
+angle $d\Omega$ and redshift interval $dz$ is
+\begin{equation}
+dV_{\rm C}= D_{\rm H}\,\frac{(1+z)^2\,D_{\rm A}^2}{E(z)}\,d\Omega\,dz
+\end{equation}
+
+ The integral of the comoving volume element
+from the present to redshift $z$ gives the total comoving volume,
+all-sky, out to redshift $z$
+
+\begin{equation}
+V_{\rm C} = \left\{
+\begin{array}{ll}
+  \left(\frac{4\pi\,D_{\rm H}^3}{2\,\Omega_k}\right)\,
+  \left[\frac{D_{\rm M}}{D_{\rm H}}\,
+  \sqrt{1+\Omega_k\,\frac{D_{\rm M}^2}{D_{\rm H}^2}}
+  -\frac{1}{\sqrt{|\Omega_k|}}\,
+  {\rm arcsinh}\left(\sqrt{|\Omega_k|}\,\frac{D_{\rm M}}{D_{\rm H}}\right)\right]
+  & {\rm for}~\Omega_k>0 \\
+  \frac{4\pi}{3}\,D_{\rm M}^3
+  & {\rm for}~\Omega_k=0 \\
+  \left(\frac{4\pi\,D_{\rm H}^3}{2\,\Omega_k}\right)\,
+  \left[\frac{D_{\rm M}}{D_{\rm H}}\,
+  \sqrt{1+\Omega_k\,\frac{D_{\rm M}^2}{D_{\rm H}^2}}
+  -\frac{1}{\sqrt{|\Omega_k|}}\,
+  {\rm arcsin}\left(\sqrt{|\Omega_k|}\,\frac{D_{\rm M}}{D_{\rm H}}\right)\right]
+  & {\rm for}~\Omega_k<0
+\end{array}
+\right.
+\end{equation}
+
+\section{COMPAS configuration fiducial values}\label{sec:appendix_COMPAS_fiducial}
+
+\begin{table*}[ht!]
+\centering
+\caption{Initial values and default settings for the \COMPAS fiducial model.}
+%
+\label{tab:app_COMPAS_fiducial}
+% ?\textwidth}{l @{\extracolsep{\fill}}
+\resizebox{\textwidth}{!}{%
+% \begin{adjustbox}{max width=1\textwidth}
+% \centering
+\begin{tabular}{lll}
+\hline  \hline
+Description and name                                 														& Value/range                       & Note / setting   \\ \hline  \hline
+\multicolumn{3}{c}{Initial conditions}                                                                      \\ \hline
+Initial primary mass \monei                               															& $[5, 150]$\Msun    & \citet{Kroupa_2001} IMF  $\propto  {\monei}^{-\alpha}$  with $\alpha_{\rm{IMF}} = 2.3$ for stars above $5$\Msun	  \\
+%
+Initial mass ratio $\qi = \mtwoi / \monei $           												& $[0, 1]$                          &       We assume a flat mass ratio distribution  $p(\qi) \propto  1$ with \mtwoi $\geq 0.1\Msun$   \\
+%
+Initial semi-major axis \ai                                            											& $[0.01, 1000]$\AU & Distributed flat-in-log $p(\ai) \propto 1 / {\ai}$ \\
+%
+Initial metallicity \Zi                                           											& $[0.0001, 0.03]$                 & Distributed uniform-in-log   \\
+%
+Initial orbital eccentricity \ei                                 							 				& 0                                & All binaries are assumed to be circular at birth  \\
+%
+% initial rotation of stars                            															& 0                                 &                  \\  \hline
+\hline
+%initial mass function (IMF) slope 	$\alpha_{\rm{IMF}}$									&  $ 2.3$		& 	 for stars above $5$\Msun from \citet{2001MNRAS.322..231K} IMF\\  \hline
+\multicolumn{3}{c}{Fiducial parameter settings:}                                                            \\ \hline
+%
+Stellar winds  for hydrogen rich stars                                   																&      \citet{Belczynski_2010a}    &   Based on {\citet{Vink_2000, Vink_2001}}, including  LBV wind mass loss with $f_{\rm{LBV}} = 1.5$   \\
+%
+Stellar winds for hydrogen-poor helium stars &  \citet{Belczynski_2010a} & Based on   {\citet{Hamann_1998}} and  {\citealt{Vink_2005}}  \\
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%% MASS TRANSFER THINGS %%%%%%%%%%%%%%%%%%%%%%%
+%
+Max transfer stability criteria & $\zeta$-prescription & Based on \citet[][]{Vigna-Gomez_2018} and references therein     \\
+%
+Mass transfer accretion rate & thermal timescale & Limited by thermal timescale for stars  \citet[][]{Hurley_2002, Vinciguerra_2020} \\
+ & Eddington-limited  & Accretion rate is Eddington-limit for compact objects  \\
+%
+Non-conservative mass loss & isotropic re-emission &  {\citet[][]{Massevitch_1975, Bhattacharya_1991, Soberman_1997}} \\
+& &  {\citet{Tauris_2006}} \\
+%
+Case BB mass transfer stability                                														& always stable         &       Based on  \citet{Tauris_2015, Tauris_2017, Vigna-Gomez_2018}         \\
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%% CE THINGS %%%%%%%%%%%%%%%%%%%%%%%
+%
+CE prescription & $\alpha-\lambda$ & Based on  \citet{Webbink_1984, deKool_1990}  \\
+%
+CE efficiency $\alpha$-parameter                     												& 1.0                               &              \\
+%
+CE $\lambda$-parameter                               													& $\lambda_{\rm{Nanjing}}$                             &        Based on \citet{Xu_2010a, Xu_2010b} and  \citet{Dominik_2012}       \\
+%
+Hertzsprung gap (HG) donor in {CE}                       														& pessimistic                       &  Defined in \citet{Dominik_2012}:  HG donors don't survive a {CE}  phase        \\
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%.    SN THINGS   %%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+{SN} natal kick magnitude \vk                          									& $[0, \infty)$\kms & Drawn from Maxwellian distribution with standard deviation $\sigma_{\rm{rms}}^{\rm{1D}}$          \\
+%
+ {SN} natal kick polar angle $\thetak$          											& $[0, \pi]$                        & $p(\thetak) = \sin(\thetak)/2$ \\
+%
+ {SN} natal kick azimuthal angle $\phi_k$                           					  	& $[0, 2\pi]$                        & Uniform $p(\phi) = 1/ (2 \pi)$   \\
+%
+ {SN} mean anomaly of the orbit                    											&     $[0, 2\pi]$                             & Uniformly distributed  \\
+ %
+Core-collapse  {SN} remnant mass prescription          									     &  delayed                     &  From \citet{Fryer_2012}, which  has no lower {BH} mass gap  \\%
+%
+USSN  remnant mass prescription          									     &  delayed                     &  From \citet{Fryer_2012}   \\%
+%
+ECSN  remnant mass presciption                        												&                                 $m_{\rm{f}} = 1.26\Msun$ &      Based on Equation~8 in \citet{Timmes_1996}          \\
+%
+Core-collapse  {SN}  velocity dispersion $\sigma_{\rm{rms}}^{\rm{1D}}$ 			& 265\kms           & 1D rms value based on              \citet{Hobbs_2005}                          \\
+%
+ USSN  and ECSN  velocity dispersion $\sigma_{\rm{rms}}^{\rm{1D}}$ 							 	& 30\kms             &            1D rms value based on e.g.,    \citet{Pfahl_2002, Podsiadlowski_2004}    \\
+%
+PISN / PPISN remnant mass prescription               											& \citet{Marchant_2019}                    &       As implemented in \citet{Stevenson_2019}      \\
+Maximum NS mass                                      & $\rm{max}_{\rm{NS}} = 2.5$\Msun & Following \citet{Fryer_2012}            \\
+Tides and rotation & & We do not include tides and/or rotation in this study\\
+Binary fraction                                      & $f_{\rm{bin}} = 0.7$ &  \\
+Solar metallicity \Zsun                             & $\rm{Z}_{\odot}\xspace = 0.0142$ & based on {\citet{Asplund_2009}} \\
+%
+%
+\hline
+\multicolumn{3}{c}{Simulation settings}                                                                     \\ \hline
+Binary population synthesis code                                      & COMPAS &       \citet{Stevenson_2017, Barrett_2018, Vigna-Gomez_2018, Neijssel_2019} \\
+& & \citet{Broekgaarden_2019, COMPAS_2022}.        \\
+\hline \hline
+\end{tabular}%
+% \end{adjustbox}
+}
+\end{table*}
+
 \end{document}

### 2024-06-25 src/scripts/data_cache.py

@@ -1,0 +1,55 @@
+import os
+from urllib.parse import urlparse
+import paths
+import requests
+from tqdm import tqdm
+import matplotlib.pyplot as plt
+import warnings
+
+__all__ = ["DataCache"]
+
+
+
+HERE = os.path.dirname(__file__)
+warnings.filterwarnings("ignore", category=UserWarning)
+plt.style.use(f"{HERE}/matplotlibrc")
+
+
+class DataCache:
+    def __init__(self, url: str):
+        self.url = url
+        if not self.exists:
+            self._download()
+
+    @property
+    def fpath(self) -> str:
+        if not hasattr(self, "_fpath"):
+            fname = urlparse(self.url).path.split("/")[-1]
+            self.__fpath = os.path.join(paths.data, fname)
+        return self.__fpath
+
+    @property
+    def exists(self) -> bool:
+        return os.path.isfile(self.fpath)
+
+    def _download(self):
+        with requests.get(self.url, stream=True) as response:
+            response.raise_for_status()
+            # Get the total file size in bytes
+            total_size = int(response.headers.get("content-length", 0))
+
+            # Download the file with a progress bar
+
+            with open(self.fpath, "wb") as file:
+                for chunk in tqdm(
+                        response.iter_content(chunk_size=8192),
+                        total=total_size // 8192,
+                        unit="KB",
+                        unit_scale=True,
+                        desc=f"Downloading {self.url}",
+                ):
+                    file.write(chunk)
+
+    @classmethod
+    def get(cls, url: str) -> str:
+        return cls(url).fpath

### 2024-06-25 src/scripts/plot_kl_distances.py

@@ -1,0 +1,28 @@
+import pandas as pd
+import numpy as np
+import paths
+from data_cache import DataCache
+import matplotlib.pyplot as plt
+
+KL_DATA = 'https://raw.githubusercontent.com/COMPAS-Surrogate/paper_data/main/kl_distance_ci_data.csv'
+
+
+def plot_kl_data():
+    data = pd.read_csv(DataCache(KL_DATA).fpath)
+    x, y, yu, yl = data.npts, data.medians, data.upper_ci_95, data.lower_ci_95
+    x, y, yu, yl = np.array(x.tolist()), y.tolist(), yu.tolist(), yl.tolist()
+    fig, axs = plt.subplots(1, 1)
+    axs.plot(x, y, label='Median')
+    axs.fill_between(x.tolist(), yl, yu, alpha=0.3, label=r'$95\%$ C.I.')
+    axs.set_ylabel('KL Distance')
+    axs.set_xlabel('Number of GP training Points')
+    leg = axs.legend()
+    leg.legend_handles[0].set_linewidth(2.0)
+    axs.set_xlim(min(data.npts), 900)
+    axs.set_ylim(bottom=0)
+    plt.tight_layout()
+    fig.savefig(paths.figures / "kl_distances.pdf", bbox_inches="tight", dpi=300)
+
+
+if __name__ == '__main__':
+    plot_kl_data()

### 2024-06-25 src/scripts/plot_pp_plot.py

@@ -1,0 +1,72 @@
+import json
+import numpy as np
+import matplotlib.pyplot as plt
+import paths
+from data_cache import DataCache
+from collections import namedtuple
+from typing import List
+from scipy.stats import binom
+from matplotlib.ticker import FixedLocator
+
+PP_DATA_URL = 'https://raw.githubusercontent.com/COMPAS-Surrogate/paper_data/main/pp_data.json'
+
+PP_DATA = namedtuple(
+    'pp_data',
+    ['p_value', 'pp', 'label']
+)
+
+CONFIDENCE_INTERVAL = [0.68, 0.95, 0.997]
+CI_ALPHA = [0.1] * len(CONFIDENCE_INTERVAL)
+
+
+def load_pp_data() -> List[PP_DATA]:
+    with open(DataCache(PP_DATA_URL).fpath, 'r') as f:
+        loaded_data = json.load(f)
+    data = [PP_DATA(**d) for d in loaded_data]
+    return data
+
+
+def plot_pp_data():
+    data = load_pp_data()
+    x = np.linspace(0, 1, len(data[0].pp))
+
+    fig, ax = plt.subplots(1, 1)
+
+    for d in data:
+        ax.plot(x, d.pp, label=d.label)
+    _add_ci_bounds(ax, x, n_posteriors=100)
+
+    ax.set_xlabel('Credible Interval (C.I.)')
+    ax.set_ylabel('Fraction of Samples in C.I.')
+    ax.set_xlim(0, 1)
+    ax.set_ylim(0, 1)
+    leg = ax.legend(
+        loc='upper left', fontsize='small', markerscale=2.5,
+        labelspacing=0.25,handlelength=1,
+        handletextpad=0.33
+    )
+    for legobj in leg.legend_handles[:4]:
+        legobj.set_linewidth(2.0)
+    ax.set_aspect('equal')
+    ax.yaxis.set_major_locator(FixedLocator([0, 0.5, 1]))
+    ax.xaxis.set_major_locator(FixedLocator([0.5, 1]))
+
+    fig.savefig(paths.figures / "pp_plot.pdf", bbox_inches="tight", dpi=300)
+
+
+def _add_ci_bounds(ax, x, n_posteriors):
+    for i, (ci, alpha) in enumerate(zip(CONFIDENCE_INTERVAL, CI_ALPHA)):
+        edge_of_bound = (1. - ci) / 2.
[…]

### 2024-06-25 src/scripts/plot_simulation_posterior.py

@@ -1,0 +1,75 @@
+import numpy as np
+from pygtc import plotGTC
+from typing import List
+import matplotlib.pyplot as plt
+from data_cache import DataCache
+import paths
+
+POSTERIOR_URL = 'https://raw.githubusercontent.com/COMPAS-Surrogate/paper_data/main/simulation_posteriors.npz'
+
+LATEX = dict(
+    mu_z=r"$\mu_z$",
+    sigma_0=r"$\sigma_0$",
+    aSF=r"$a_{\rm SF}$",
+    dSF=r"$d_{\rm SF}$",
+)
+
+
+class PosteriorDatasets:
+    def __init__(self, posteriors: List[np.ndarray], truths: np.ndarray, params: List[str], labels: List[str]):
+        self.chains = posteriors
+        self.truths = truths
+        self.paramNames = [LATEX[p] for p in params]
+        self.chainLabels = labels
+        self.n_post, self.n_samp, self.n_dim = (len(posteriors), *posteriors[0].shape)
+        assert len(truths) == self.n_dim
+        assert len(labels) == self.n_post
+
+    @property
+    def posterior_matrix(self):
+        return np.array(self.chains)
+
+    @property
+    def paramRanges(self):
+        ranges = np.zeros((self.n_dim, 2))
+        for d in range(self.n_dim):
+            p = self.posterior_matrix[:, :, d].flatten()
+            ranges[d] = np.array([min(p), max(p)])
+        return ranges.tolist()
+
+    @classmethod
+    def from_npz(cls, fname):
+        data = np.load(fname)
+        posteriors = [p for p in data["posteriors"]]
+        return cls(posteriors, data["truths"], data["params"], data["labels"])
+
+    @classmethod
+    def from_cache(cls):
+        return cls.from_npz(DataCache(POSTERIOR_URL).fpath)
+
+    def __dict__(self):
+        return {
+            "chains": self.chains[0:2],
+            "chainLabels": self.chainLabels[0:2],
+            "truths": self.truths,
+            "paramNames": self.paramNames,
+            "paramRanges": self.paramRanges,
+            "reference": self.chains[2],
+        }
+
+if __name__ == '__main__':
+    data = PosteriorDatasets.from_cache()
+    fig = plotGTC(
+        **data.__dict__(),
+        filledPlots=True,
+        figureSize="MNRAS_page",
+        truthLabels="Truth",
+        legendMarker="All",
[…]

### 2024-06-25 src/scripts/pygtc.py

@@ -1,0 +1,1219 @@
+from packaging import version
+import warnings
+import sys
+from matplotlib import pyplot as plt
+from matplotlib.colors import LinearSegmentedColormap
+import matplotlib as mpl
+import matplotlib.colors as mplcolors
+import numpy as np
+import matplotlib.ticker as mtik
+import types
+try:
+    import scipy.ndimage
+    from scipy.stats import norm
+    haveScipy = True
+except ImportError:
+    haveScipy = False
+
+_PYVER = sys.version_info[0]
+_MPLVER = version.parse(mpl.__version__)
+
+_SNAKE_LEGEND_HANDLES = _MPLVER >= version.parse('3.6')
+
+__all__ = ['plotGTC']
+
+
+# Create a full GTC
+
+REF_COLOR = 'lightgray'
+
+def plotGTC(chains, **kwargs):
+    r"""Make a great looking Giant Triangle Confusogram (GTC) with one line of
+    code! A GTC is like a triangle (or corner) plot, but you get to put as
+    many sets of data, and overlay as many truths as you like. That's what can
+    make it so *confusing*!
+
+    Parameters
+    ----------
+    chains : array-like[nSamples,nDims] or a
+            list[[nSamples1,nDims], [nSamples2,nDims], ...]
+        All chains (where a chain is [nSamples,nDims]) in the list must have
+        the same number of dimensions. Note: If you are using ``emcee``
+        (http://dan.iel.fm/emcee/current/) - and you should! - each element
+        of chains is an ``EnsembleSampler.flatchain`` object.
+
+    Keyword Arguments
+    -----------------
+    weights : array-like[nSamples] or a list[[nSamples1], ...]
+        Weights for the sample points. The number of 1d arrays passed must
+        correspond to the number of `chains`, and each `weights` array must
+        have the same length nSamples as its corresponding chain.
+
+    chainLabels : array-like[nChains]
+        A list of text labels describing each chain passed to chains.
+        len(chainLabels) must equal len(chains). chainLabels supports LaTex
+        commands enclosed in $..$. Additionally, you can pass None as a label.
+        Default is ``None``.
+
+    paramNames : list-like[nDims]
+        A list of text labels describing each dimension of chains.
+        len(paramNames) must equal nDims=chains[0].shape[1]. paramNames
+        supports LaTex commands enclosed in $..$. Additionally, you can pass
+        None as a label. Default is None, however if you pass a
[…]

### 2024-06-25 src/scripts/plot_kl_distances.py
add morre plots

@@ -14,7 +14,7 @@ def plot_kl_data():
     fig, axs = plt.subplots(1, 1)
     axs.plot(x, y, label='Median')
     axs.fill_between(x.tolist(), yl, yu, alpha=0.3, label=r'$95\%$ C.I.')
-    axs.set_ylabel('KL Distance')
+    axs.set_ylabel('JS divergence')
     axs.set_xlabel('Number of GP training Points')
     leg = axs.legend()
     leg.legend_handles[0].set_linewidth(2.0)

### 2024-06-25 src/scripts/plot_ogc4_events.py

@@ -1,0 +1,10 @@
+from ogc4_interface.population_mcz import PopulationMcZ
+import paths
+
+if __name__ == '__main__':
+    pop = PopulationMcZ.load()
+    fig, axes = pop.plot_event_mcz_estimates()
+    # remve axes minor ticks along y
+    for ax in fig.get_axes():
+        ax.tick_params(axis='y', which='minor', length=0)
+    fig.savefig(paths.figures / "ogc4_events.pdf", bbox_inches="tight", dpi=70)

### 2024-06-25 src/scripts/plot_ogc4_weights.py

@@ -1,0 +1,8 @@
+from ogc4_interface.population_mcz import PopulationMcZ
+import paths
+
+if __name__ == '__main__':
+    pop = PopulationMcZ.load()
+    pop = pop.filter_events(threshold=0.95)
+    fig = pop.plot_weights().get_figure()
+    fig.savefig(paths.figures / "ogc4_weights.pdf", bbox_inches="tight", dpi=300)

### 2024-06-25 src/scripts/plot_simulation_posterior.py

@@ -3,6 +3,7 @@ from pygtc import plotGTC
 from typing import List
 import matplotlib.pyplot as plt
 from data_cache import DataCache
+from scipy.stats import entropy
 import paths
 
 POSTERIOR_URL = 'https://raw.githubusercontent.com/COMPAS-Surrogate/paper_data/main/simulation_posteriors.npz'
@@ -15,6 +16,25 @@ LATEX = dict(
 )
 
 
+
+def compute_kl_div(res1_posterior, res2_posterior, use_kde=False, bandwidth=0.2, bins=100):
+    assert res1_posterior.shape[1] == res2_posterior.shape[1], "Dimensions of posteriors do not match."
+    js_divs, ks_stats, kl_divs = [], [], []
+    ndim = res1_posterior.shape[1]
+    for d in range(ndim):
+        p, q = res1_posterior[:,d], res2_posterior[:,d]
+        x_grid = np.linspace(min(p.min(), q.min()), max(p.max(), q.max()), 1000)
+        min_len = min(len(p), len(q))
+        p, q = np.random.choice(p, min_len), np.random.choice(q, min_len)
+        p_density, _ = np.histogram(p, bins=bins, range=(x_grid.min(), x_grid.max()), density=True)
+        q_density, _ = np.histogram(q, bins=bins, range=(x_grid.min(), x_grid.max()), density=True)
+        p_density += 1e-10
+        q_density += 1e-10
+        p_density /= p_density.sum()
+        q_density /= q_density.sum()
+        kl_divs.append(entropy(p_density, q_density))
+    return f"{np.mean(kl_divs):.2f}"
+
 class PosteriorDatasets:
     def __init__(self, posteriors: List[np.ndarray], truths: np.ndarray, params: List[str], labels: List[str]):
         self.chains = posteriors
@@ -22,8 +42,13 @@ class PosteriorDatasets:
         self.paramNames = [LATEX[p] for p in params]
         self.chainLabels = labels
         self.n_post, self.n_samp, self.n_dim = (len(posteriors), *posteriors[0].shape)
+        kl_1 = compute_kl_div(posteriors[0], posteriors[2])
+        kl_2 = compute_kl_div(posteriors[1], posteriors[2])
+        self.chainLabels[0]+= f" ($D={kl_1}$)"
+        self.chainLabels[1] += f" ($D={kl_2}$)"
         assert len(truths) == self.n_dim
         assert len(labels) == self.n_post
+
 
     @property
     def posterior_matrix(self):

### 2024-06-26 src/scripts/plot_simulation_posterior.py
pushing changes

@@ -1,7 +1,6 @@
 import numpy as np
 from pygtc import plotGTC
 from typing import List
-import matplotlib.pyplot as plt
 from data_cache import DataCache
 from scipy.stats import entropy
 import paths
@@ -16,13 +15,12 @@ LATEX = dict(
 )
 
 
-
 def compute_kl_div(res1_posterior, res2_posterior, use_kde=False, bandwidth=0.2, bins=100):
     assert res1_posterior.shape[1] == res2_posterior.shape[1], "Dimensions of posteriors do not match."
     js_divs, ks_stats, kl_divs = [], [], []
     ndim = res1_posterior.shape[1]
     for d in range(ndim):
-        p, q = res1_posterior[:,d], res2_posterior[:,d]
+        p, q = res1_posterior[:, d], res2_posterior[:, d]
         x_grid = np.linspace(min(p.min(), q.min()), max(p.max(), q.max()), 1000)
         min_len = min(len(p), len(q))
         p, q = np.random.choice(p, min_len), np.random.choice(q, min_len)
@@ -35,6 +33,7 @@ def compute_kl_div(res1_posterior, res2_posterior, use_kde=False, bandwidth=0.2,
         kl_divs.append(entropy(p_density, q_density))
     return f"{np.mean(kl_divs):.2f}"
 
+
 class PosteriorDatasets:
     def __init__(self, posteriors: List[np.ndarray], truths: np.ndarray, params: List[str], labels: List[str]):
         self.chains = posteriors
@@ -44,11 +43,10 @@ class PosteriorDatasets:
         self.n_post, self.n_samp, self.n_dim = (len(posteriors), *posteriors[0].shape)
         kl_1 = compute_kl_div(posteriors[0], posteriors[2])
         kl_2 = compute_kl_div(posteriors[1], posteriors[2])
-        self.chainLabels[0]+= f" ($D={kl_1}$)"
+        self.chainLabels[0] += f" ($D={kl_1}$)"
         self.chainLabels[1] += f" ($D={kl_2}$)"
         assert len(truths) == self.n_dim
         assert len(labels) == self.n_post
-
 
     @property
     def posterior_matrix(self):
@@ -81,6 +79,7 @@ class PosteriorDatasets:
             "paramRanges": self.paramRanges,
             "reference": self.chains[2],
         }
+
 
 if __name__ == '__main__':
     data = PosteriorDatasets.from_cache()

### 2024-06-26 src/scripts/plot_simulation_posterior.py
fix path

@@ -96,4 +96,4 @@ if __name__ == '__main__':
     # turn off all minor ticks
     for ax in fig.axes:
         ax.tick_params(which='minor', length=0)
-    fig.savefig(paths.figures / "simulation_posteriors.pdf", bbox_inches="tight", dpi=300)
+    fig.savefig(paths.figures / "simulation_posterior.pdf", bbox_inches="tight", dpi=300)

### 2024-06-26 src/tex/ms.tex
fix latex error
add extra plot
fix path
update tex

@@
 
+\usepackage{float}
+\makeatletter
+\let\newfloat\newfloat@ltx
+\makeatother
+
+\usepackage{algorithm}
+\usepackage{algorithmic}
+
+\usepackage{tikz}
+\usetikzlibrary{shapes.geometric, arrows}
+
 \setlength{\marginparwidth}{1.5cm}
@@
 
-\input{custom_commands}
+\newcommand\COMPAS{{\sc{COMPAS}}\xspace}
+\newcommand*\diff{\mathop{}\!\mathrm{d}}
+\newcommand{\MSFR}{\ensuremath{{M}_{\rm{SFR}}}\xspace}
+\newcommand{\ts}{\ensuremath{{t}_{\rm{s}}}\xspace}
+\newcommand{\tsup}{\textsuperscript}
+\newcommand{\Vc}{\ensuremath{{V}_{\rm{c}}}\xspace}
+
+\newcommand{\monei}{\ensuremath{m_{1,\rm{i}}}\xspace}
+\newcommand{\mtwoi}{\ensuremath{m_{2,\rm{i}}}\xspace}
+\newcommand{\monef}{\ensuremath{m_{1,\rm{f}}}\xspace}
+\newcommand{\mtwof}{\ensuremath{m_{2,\rm{f}}}\xspace}
+\newcommand{\ai}{\ensuremath{a_{\rm{i}}}\xspace}
+\newcommand{\qi}{\ensuremath{q_{\rm{i}}}\xspace}
+\newcommand{\Zi}{\ensuremath{Z_{\rm{i}}}\xspace}
+\newcommand{\vk}{\ensuremath{v_{\rm{k}}}\xspace}
+\newcommand{\thetak}{\ensuremath{{\theta}_{\rm{k}}}\xspace}
+\newcommand{\phik}{\ensuremath{{\phi}_{\rm{k}}}\xspace}
+\newcommand{\ei}{\ensuremath{{e}_{\rm{i}}}\xspace}
+
+\newcommand{\Msun}{\ensuremath{\,\rm{M}_{\odot}}\xspace}
+\newcommand{\Zsun}{\ensuremath{\,\rm{Z}_{\odot}}\xspace}
+\newcommand{\kms}{\ensuremath{\,\rm{km}\,\rm{s}^{-1}}\xspace}
+\newcommand{\AU}{\ensuremath{\,\mathrm{AU}}\xspace}
+\newcommand{\Mc}{\ensuremath{\mathcal{M}}\xspace}
+
+\newcommand{\Lmain}{\ensuremath{\mathcal{L}(\mathcal{D}|\lambda)}\xspace}
+\newcommand{\Lp}{\ensuremath{\mathcal{L}_{\rm p}}\xspace}
+\newcommand{\Le}{\ensuremath{\mathcal{L}_{\rm event}}\xspace}
+\newcommand{\pg}{\ensuremath{p_{\rm grid}}\xspace}
+\newcommand{\Lsurr}{\ensuremath{\mathcal{L}^{\star}(\lambda)}\xspace}
+
+%%%%%%%%%%%%%%% SYMBOLS
+
+% Projects:
+\newcommand{\project}[1]{\textsf{#1}}
+
+\newcommand{\python}{\project{Python}}
+\newcommand{\jupyter}{\project{Jupyter}}
+
+\newcommand{\exoplanet}{\project{exoplanet}}
+\newcommand{\lightkurve}{\project{lightkurve}}
+\newcommand{\starry}{\project{starry}}
+\newcommand{\pymc}{\project{PyMC3}}
+\newcommand{\pymcextra}{\project{pymc3-ext}}
+\newcommand{\celerite}{\project{celerite}}
+\newcommand{\astropy}{\project{astropy}}
+\newcommand{\scipy}{\project{scipy}}
+\newcommand{\jupytext}{\project{jupytext}}
+\newcommand{\sphinx}{\project{sphinx}}
+\newcommand{\jupyterbook}{\project{Jupyter-book}}
+\newcommand{\arviz}{\project{ArviZ}}
+\newcommand{\nbconvert}{\project{nbconvert}}
+\newcommand{\numpy}{\project{numpy}}
+\newcommand{\pandas}{\project{pandas}}
+\newcommand{\matplotlib}{\project{matplotlib}}
+\newcommand{\corner}{\project{corner}}
+\newcommand{\emcee}{\project{emcee}}
+\newcommand{\trieste}{\project{trieste}}
+\newcommand{\tensorflow}{\project{TensorFlow}}
+\newcommand{\pycbc}{\project{PyCBC}}
+
+\newcommand{\LVK}{\project{LVK}}
+\newcommand{\mast}{\project{MAST}}
+\newcommand{\exofop}{\project{ExoFOP}}
+\newcommand{\tessAtlas}{\project{TESS Atlas}}
+\newcommand{\ogc}{\project{4-OGC}}
+
+% math
+\newcommand{\T}{\ensuremath{\mathrm{T}}}
+\newcommand{\dd}{\ensuremath{ \mathrm{d}}}
+% \newcommand{\unit}[1]{{\ensuremath{ \mathrm{#1}}}}
+\newcommand{\bvec}[1]{{\ensuremath{\boldsymbol{#1}}}}
+
+\DeclareMathOperator{\invG}{Inv-\mathnormal{\Gamma}}
+\DeclareMathOperator{\N}{\mathcal{N}}
+\DeclareMathOperator{\U}{\mathcal{U}}
+\DeclareMathOperator{\Un}{\mathcal{U}}
+\DeclareMathOperator{\Par}{\mathcal{P}ar}
+\DeclareMathOperator{\tmin}{\mathnormal{t_{\rm min}}}
+\DeclareMathOperator{\tmax}{\mathnormal{t_{\rm max}}}
+
+%% affiliation shortcuts
+\newcommand{\SPA}{School of Physics and Astronomy, Monash University, Clayton VIC 3800, Australia}
+\newcommand{\OzGravMonash}{OzGrav: The ARC Centre of Excellence for Gravitational Wave Discovery, Clayton VIC 3800, Australia}
+\newcommand{\AMNH}{Department of Astrophysics, American Museum of Natural History, New York, NY 10024, USA}
+\newcommand{\CCA}{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}
+\newcommand{\CUNY}{Graduate Center, City University of New York, 365 5th Avenue, New York, NY 10016, USA}
+\newcommand{\BMCC}{Department of Science, BMCC, City University of New York, New York, NY 10007, USA}
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%
+
+\definecolor{orange-red}{rgb}{1.0, 0.27, 0.0}
+\newcommand{\jeff}[1]{\textbf{\textcolor{orange-red}{#1}}}
+\newcommand{\ilya}[1]{\textbf{\textcolor{magenta}{#1}}}
+\newcommand{\avi}[1]{\textbf{\textcolor{cyan}{#1}}}
+\newcommand{\chayan}[1]{\textbf{\textcolor{green}{#1}}}
+\newcommand{\resp}[1]{#1}
 
@@
 \label{sec:intro}
-Over the past decade, the improvements in detector sensitivity have led to an exponential increase in the rate of GW detections.
-With next-generation detectors on the horizon (such as CE, ET, LISA), the number of GW detections will increase even more.
+Over the past decade, improvements in detector sensitivity have led to an exponential increase in the rate of GW detections.
+With next-generation detectors on the horizon (such as CE, ET, and LISA), the number of GW detections will increase even more.
 The most recent catalogue of GW contains
@@
 \section{Method}
-\subsection{Gaussian process and Bayesian Optimisation}
-Talk about bayesian optimisation
 
@@
 \subsection{Bayesian Model}
-We define the likelihood $\mathcal{L}$ that we will see the data $\mathcal{D}$ for a particular set of parameters $\lambda$ as
-
+Given $N_{\mathrm{obs}}$ observations from data  $\mathcal{D} \equiv \{D_1, D_2, ...
+D_{N_{\mathrm{obs}}}\}$, we define the likelihood $\mathcal{L}$ that we will see data $\mathcal{D}$ for a particular set of parameters $\lambda$ as
 \begin{equation}
-  \log{\mathcal{L}\left( \mathcal{D} |\lambda\right)} = \log{\Lp\left(N_{\mathrm{obs}}|\lambda\right)} +
-  \sum_{i=1}^{N_\mathrm{obs}} \log{\Le(D_i|\lambda)},
+  \log{\Lmain} = \log{\Lp\left(N_{\mathrm{obs}}|\lambda\right)} +
+  \sum_{i=1}^{N_\mathrm{obs}} \log{\Le(D_i|\lambda)}\ .
   \label{eq:methods_likelihood}
@@
 
-\bigskip\noindent
-where $N_{\mathrm{obs}}$ is the number of observations in the data $\mathcal{D} \equiv \{D_1, D_2, ... D_{N_{\mathrm{obs}}}\}$.
-
+\bigskip
 The first term of Equation~\ref{eq:methods_likelihood} is a Poisson likelihood on there being $N_\mathrm{obs}$ detections,
@@
 where $\pg(z=z_i,\mathcal{M}=\mathcal{M}_{i}|\lambda)$ is the entry in the probability distribution matrix $\pg(z,\mathcal{M}|\lambda)$ in the $z$ and $\mathcal{M}$ bin of the observed $i^{th}$ event.
+
 The probability distribution matrix $\pg(z,\mathcal{M}|\lambda)$ is constructed by dividing the entries in the matrix of expected detections for a given $\lambda$, generated by multiplying the surrogate model detection rate matrix by the run duration, by the sum of the entries in the matrix, $\mu$.
 
-\subsection{Using gravitational wave observations}\label{sec:method_LVK}
-
-The catalogue of LIGO, Virgo, and KAGRA (LVK) gravitational wave observations (\citet{GWTC-2_1_zenodo}, \citet{GWTC-3_zenodo}) provides us with a number of merger detections from a population of \acp{DCO}.
-For each LVK event we know, among other things, the chirp mass of the merging binary and the merger redshift, up to a measurement uncertainty.
-We only include confident detections with a minimum astrophysical probability $p_{astro} \ge 0.95$; although less confident events could be included \citep[e.g.][]{Farr_2015}, they will typically contribute little information due to greater measurement uncertainties, so we omit them for this proof-of-concept study.
-
-Once we validate our method, we can use our surrogate model and the LVK data to infer real physical constraints on the parameters under study, with the important caveat that we have fixed the astrophysical parameters, and any errors in those could bias the inferred \ac{MSSFR} model values.
-
-\subsubsection{LVK likelihood calculation}\label{sec:method_LVK_likelihood_calculation}
-
- In practice, for LVK events, we may not know the values $z_i, \Mc_{i}$ and hence the probability $\Le(z=z_i,\Mc=\Mc_{i}|\lambda)$ for the observed $i^{th}$ event in Equation~\ref{eq:methods_likelihood_term2} perfectly, but may only have $K$ samples from a posterior $p(z_i,\Mc_{i}|D_i)$.
-For LVK data, we substitute the following for Equation~\ref{eq:methods_likelihood_term2}:
+ In practice, we may not know the values $z_i, \Mc_{i}$ perfectly,  but may only have $K$ samples from a posterior $p(z_i,\Mc_{i}|D_i)$.
+ Hence the probability $\Le(z=z_i,\Mc=\Mc_{i}|\lambda)$ for the  $i^{th}$ event in Equation~\ref{eq:methods_likelihood_term2} can be re-written as
 
@@
 \bigskip\noindent
-where the subscript $k$ refers to the $k^{th}$ posterior sample among $K$ available samples for event $i$ and $\pi(z,\Mc)$ is the prior used by LVK data analysts in the original inference on the observed events \citep[see, e.g.][]{Mandel_2019}.
+where the subscript $k$ refers to the $k^{th}$ posterior sample among $K$ available samples, and $\pi(z,\Mc)$ is the prior used in the original inference for the event~\citep[see, e.g.][]{Mandel_2019}.
 
-In practice, if the number of bins where an event has posterior support is much smaller than the number of samples $K$, it may be more efficient to pre-compute the weight that each observed event $i$ contributes to each of the bins $b$, $w_{i,b}$, which is independent of $\lambda$:
+ If the number of bins where an event has posterior support is much smaller than the number of samples $K$, it may be more efficient to pre-compute the weight that each observed event $i$ contributes to each of the bins $b$, $w_{i,b}$, which is independent of $\lambda$:
 
@@
 \begin{equation}
-    \Le(D_i|\lambda) = \sum_b \pg(z^b, \Mc^b | \lambda) w_{i,b}
+    \Le(D_i|\lambda) = \sum_b \pg(z^b, \Mc^b | \lambda)\,  w_{i,b}\, .
 \end{equation}
 
+Unfortunatly, \Lmain is an expensive function, making it hard to sample with.
+In the next section, we discuss a surrogate model for \Lmain that is cheaper to sample with.
+
+\subsection{GP surrogate using Bayesian optimisation }
+Given  $M$ ``observations" of \Lmain, $\{\mathcal{L}(\lambda_0), ..., \mathcal{L}(\lambda_M)\}$, we  define a conditional probabilty distribution
+\begin{align}
+    \Lsurr &= \mathcal{GP}( \mathcal{L}(\lambda) | \mathcal{L}(\lambda_0), ..., \mathcal{L}(\lambda_M))\, \nonumber \\
+    &= \mathcal{GP}(\mu(\lambda), \Sigma(\lambda))
+\end{align}
+where the conditional probability distribution is given by a Guassian process $\mathcal{GP}$, of mean $\mu$ and covariane $\Sigma$.
+A Gaussian process, provides a Bayesian posterior probability
+distribution that describes potential values for $\Lmain$ at a candidate point $\lambda$.
+Each time we observe a new $\mathcal{L}(\lambda')$, this posterior distribution is updated.
+
+\begin{algorithm}
+\caption{Basic pseudo-code for Bayesian optimization}
+\label{alg:BayesOpt}
+\begin{algorithmic}[1] % [1] turns on line numbering
+    \STATE Place Gaussian process $\mathcal{GP}$ prior on $\Lmain$
+    \STATE Observe $\Lmain$ at $N_0$ points randomly sampled.
+    \WHILE{$n \le N$}
+        \STATE Update the posterior probability distribution on $\Lmain$ using all available data
+        \STATE Let $\lambda_{\rm best}$ be a maximizer of the acquisition function over $\lambda$, where the acquisition function is computed using the current posterior distribution.
+        \STATE Observe $L_{\rm best} = \mathcal{L}(\mathcal{D}|\lambda_{\rm best})$.
+        \STATE Increment $n$
+    \ENDWHILE
+    \STATE Return a solution: $\Lsurr \sim \mathcal{GP}(\lambda)$
+\end{algorithmic}
+\end{algorithm}
+
+The acquisition function measures the value that would be generated by
+evaluation of the objective function at a new point $\lambda'$, based on the current posterior distribution over \Lmain. We discuss expected improvement, the most commonly used ac
+
+Bayesian optimisation modeling using Gaussian processes is discussed in detail in~\cite{}.
+
+\tikzstyle{startstop} = [rectangle, rounded corners,
+minimum width=3cm,
+minimum height=1cm,
+text centered,
+draw=black,
+fill=red!30]
+
+\tikzstyle{process} = [rectangle,
+minimum width=3cm,
+minimum height=1cm,
+text centered,
+text width=3cm,
+draw=black,
+fill=orange!30]
+
+\tikzstyle{decision} = [diamond,
+minimum width=3cm,
+minimum height=1cm,
+text centered,
+draw=black,
+fill=green!30]
+\tikzstyle{arrow} = [thick,->,>=stealth]
+
+\begin{tikzpicture}[node distance=2cm]
+
+\node (start) [startstop] {Sample $\pi(\lambda)$};
+\node (compute) [process, below of=start] {Compute $\Lmain$};
+\node (train) [process, below of=compute] {Train \Lsurr};
+\node (sample) [process, below of=train] {Sample
+$p(\lambda|\mathcal{D})\sim\Lsurr\pi(\lambda)$};
+\node (check) [decision, below of=sample, yshift=-0.5cm] {\mathcal{O}(\lambda) > N?};
+
+\node (stop) [startstop, below of=check] {Stop};
+\node (acquire) [process, right of=check, xshift=2cm] {Acquire $\lambda$};
+
+\draw [arrow] (start) -- (compute);
+\draw [arrow] (compute) -- (train);
+\draw [arrow] (train) -- (sample);
+\draw [arrow] (sample) -- (check);
+\draw [arrow] (check) -- node[anchor=east] {no} (acquire);
+\draw [arrow] (check) -- node[anchor=south] {yes} (stop);
+\draw [arrow] (acquire) |- (compute);
+
+\end{tikzpicture}
+
+\subsection{Acquisition Functions}
+
+An acquisition function is a function that maps the posterior predictive mean and variance to a scalar value. The scalar value is used to select the next point to observe from the model.
+
+In general, we will write the acquisition function as $\alpha(\mu,\sigma)$, where $\mu$ is the posterior predictive mean and $\sigma$ is the posterior predictive variance.
+
+Each criterion performs a trade-off between exploration i.e. investigating regions where the variance is large and exploitation i.e. investigating regions where the prediction is minimized.
+
+\paragraph{Uncertainty-based exploration (UE)}
+
+This acquisition function aims at reducing the uncertainty of the GP model. It is defined as
+
+$$
+\alpha_{\rm{UE}}^i(\sigma^i)= \sigma^i
+$$
+
+and helps to explore the function space in regions of high uncertainty.
+
+One of the most popular BO algorithms is the Efficient Global Optimization (EGO) \cite{jones1998efficient}.
+It uses GP as surrogate model and the Expected Improvement as the infill sampling criterion. This work is focused on BO algorithms using GP and its variants as surrogate models.
+
+\paragraph{Expected Improvement}
+
+This also focuses on the maximisation/minimisation of the GP's posterior predictive mean.
+It looks for the best value of $\mu$ so far and then looks for the next point that has a higher probability of being better than the best value so far.
+
+$$
+\alpha_{\rm{EI}}^i(\mu^i, \sigma^i)= \sigma^i\ (u^i \mathcal{N}_{\rm CDF}(u^i) \mathcal{N}_{\rm PDF}(u^i)\ ,
+$$
+
+where $u^i = \frac{\mu^i - \mu_{\rm{best}}}{\sigma^i}$ and $\mu_{\rm{best}}$ is the best value of $\mu$ so far.
+
+where $\mathcal{N}_{\rm PDF}$ denotes the Probability Density function (PDF) of the univariate Gaussian probability distribution.
+
+While we would like to choose $x$ so that this improvement is large, $f(x)$ is unknown until after the evaluation.  What we can do, however,
+is to take the expected value of this improvement and choose $x$ to maximize it.
+
+The Expected Improvement (EI)takes into account the improvement induced by a candidate that is defined as: $I(\textbf{x})=\text{max}\{0,y_{\text{min}}-f(\textbf{x})\}$.
+
 \subsection{Diagnostics}
-variable LnL and Kl divergence
+
+\begin{equation}
+    \mathcal{L}^{\star}(\lambda) = \mu_{\mathcal{GP}}(\lambda)
+\end{equation}
+\begin{equation}
+    \mathcal{L}^{\star}_{V}(\lambda) = \mathcal{N}(\mu_{\mathcal{GP}}(\lambda), \sigma_{\mathcal{GP}}(\lambda))
+\end{equation}
 
 %--------------------------------------------------------
-\section{Toy model}
+\section{Inference with surrogate model}
+
+\subsection{Simulation study}
+
+To validate our method, following Section 2.5 of \citep{}, we sample 100 $\lambda$ from a $\pi(\lambda)$, and generate $\vec{\mathcal{D}}$, unique mock \COMPAS populations for each $\lambda$.
+We build surrogate likelihoods for each dataset, and use them to approximate $p(\lambda|\mathcal{D})$.
+Figure~\ref{fig:simulation_posterior} displays the $p(\lambda|\mathcal{D})$ for one mock population.
+The gray ``reference'' posterior is obtained using a surrogate likelihood $\mathcal{L}^{\star}$ tranined with 950 points.
+The blue posterior is obtained using a surrogate likelihood $\mathcal{L}^{\star}_V$ trained with 50 points, while the purple posterior's surrogate is built with 325 points.
+The blue posterior has a $D_{KL}>10$ when compared with the gray reference posterior.
+The purple  posterior has a $D_{KL}<1$ when compared with the gray reference posterior, indicating that the two posteriors are similar.
+The orange line marks the true $\lambda$ used to generate the dataset.
+\begin{figure}[ht!]
+    \script{plot_simulation_posterior.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/simulation_posterior.pdf}
+        \caption{
+            Simulation posteriors
+        }
+        \label{fig:simulation_posterior}
+    \end{centering}
+\end{figure}
+
+The PP-plot in Figure~\ref{fig:pp_plot} shows that the true value lies withing the correct region of the credible intervals, indicating that the results are not biased.
+\begin{figure}[ht!]
+    \script{plot_pp_plot.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/pp_plot.pdf}
+        \caption{
+            PP-plots from simulation study.
+        }
+        \label{fig:pp_plot}
+    \end{centering}
+\end{figure}
+
+Finally, Figure~\ref{} shows the JS diveregence
+
+The catalogue of LIGO, Virgo, and KAGRA (LVK) gravitational wave observations (\citet{GWTC-2_1_zenodo}, \citet{GWTC-3_zenodo}) provides us with a number of merger detections from a population of \acp{DCO}.
+For each LVK event we know, among other things, the chirp mass of the merging binary and the merger redshift, up to a measurement uncertainty.
+We only include confident detections with a minimum astrophysical probability $p_{astro} \ge 0.95$; although less confident events could be included \citep[e.g.][]{Farr_2015}, they will typically contribute little information due to greater measurement uncertainties, so we omit them for this proof-of-concept study.
+
+Once we validate our method, we can use our surrogate model and the LVK data to infer real physical constraints on the parameters under study, with the important caveat that we have fixed the astrophysical parameters, and any errors in those could bias the inferred \ac{MSSFR} model values.
+
 Include corners of variable LnL + normal LnL
@@
 
-\begin{figure}[ht!]
-    \script{plot_pp_plot.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/pp_plot.pdf}
-        \caption{
-            PP-plots from simulation study.
-        }
-        \label{fig:pp_plot}
-    \end{centering}
-\end{figure}
-
 %--------------------------------------------------------
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/ogc4_events.pdf}
+        \includegraphics[width=0.8\columnwidth]{figures/ogc4_events.pdf}
         \caption{
@@
 
+We do not
+\begin{itemize}
+    \item Do hyper-parameter optimisation for the GP (kernal choices, kernel settings)
+    \item Do hyper-parameter optimisation for the acquisition function
+    \item Explore range of acquisition functions
+\end{itemize}
+
 %%% DATA AVAILABILITY %%%
@@
 \facilities{LIGO-VIRGO-KAGRA}
-
+%
 \software{
@@
 \astropy~\citep{astropy},
-\arviz~\citep{arviz_2019},
-\exoplanet~\citep{Foreman-Mackey:2021:JOSS},
-\lightkurve~\citep{LightkurveCollaboration:2018:ascl},
-\starry~\citep{Luger:2019:AJ},
-\celerite~\citep{Foreman-Mackey:2017:ascl},
-\pymc~\citep{Salvatier:2016:ascl},
+\COMPAS~\citep{COMPAS_SOFTWARE_2022},
+\pycbc~\citep{},
+\emcee~\citep{},
+\trieste~\citep{},
+\tensorflow~\citep{},
 \numpy~\citep{numpy},
@@
 \sphinx~\citep{sphinx_doc},
-\jupyter~\citep{jupyter},
-\jupyterbook~\citep{jupyter_book}.
+\jupyterbook~\citep{jupyter,jupyter_book}.
 }

### 2024-06-26 src/tex/ms.tex
[showyourwork]
[showyourwork] overleaf sync

@@
 $p(\lambda|\mathcal{D})\sim\Lsurr\pi(\lambda)$};
-\node (check) [decision, below of=sample, yshift=-0.5cm] {\mathcal{O}(\lambda) > N?};
+\node (check) [decision, below of=sample, yshift=-0.5cm] {$\mathcal{O}(\lambda) > N?$};
 
@@
 %--------------------------------------------------------
-\section{LVK analysis}
-\subsection{Data}
+\subsection{LVK analysis}
+\paragraph{}{Data}
 
-\begin{figure}[ht!]
-    \script{plot_ogc4_events.py}
-    \begin{centering}
-        \includegraphics[width=0.8\columnwidth]{figures/ogc4_events.pdf}
-        \caption{
-            The events and uncertainty as predicted by Ogc4
-        }
-        \label{fig:ogc4_events}
-    \end{centering}
-\end{figure}
+We get the data from the \ogc.
+Plots of the event $\Mc,z$ median estimates and $X-\sigma$ uncertainty can be found in Figure X in the Appendix.
 
@@
 \appendix
+
+\section{OGC4 Events}
+
+\begin{figure}[ht!]
+    \script{plot_ogc4_events.py}
+    \begin{centering}
+        \includegraphics[width=0.8\columnwidth]{figures/ogc4_events.pdf}
+        \caption{
+            The events and uncertainty as predicted by Ogc4
+        }
+        \label{fig:ogc4_events}
+    \end{centering}
+\end{figure}
 

### 2024-06-26 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-We build upon Riley at al. (2023)'s work, and
+We build upon \citep{Riley:2023:ApJ}'s work, and
 
@@
 \subsection{MSSFR parameters}\label{sec:method_parameters}
-
+\avi{SECTION TAKEN FROM \citep{Riley:2023:ApJ}, FIGURE OUT OUT HOW CONDENSE}
 The parameters we chose to vary for this study are four free parameters for the calculation of the \ac{MSSFR} in the phenomenological model of \citet{Neijssel_2019}.
@@
 \subsection{Bayesian Model}
+We use the Bayesian model described in \citep{Riley:2023:ApJ}.
 Given $N_{\mathrm{obs}}$ observations from data  $\mathcal{D} \equiv \{D_1, D_2, ...
@@
 We get the data from the \ogc.
-Plots of the event $\Mc,z$ median estimates and $X-\sigma$ uncertainty can be found in Figure X in the Appendix.
+Plots of the event $\Mc,z$ median estimates and $X-\sigma$ uncertainty can be found in Figure~\ref{fig:ogc4_events} in the Appendix~\ref{}. 
+Filtering the events based on the OGC definition of $p(bbh)\geq0.95$ and ensuring that the median value of the $p(\Mc,z|d_i)$ lies within $\Mc\in[4,45]$, $z\in[0,1]$, we get XX events. 
+These are colored green in Fig~\ref{fig:ogc4_events}.
+
+Using Equation~\ref{}, we obtain the weights for each posterior. 
+A plot of each event's weights summed together is shown in Figure~\ref{fig:ogc4_weights}.
 
@@
 
-We do not 
+Future work
 \begin{itemize}
@@
 
-\section{OGC4 Events}
+\section{Interfacing with OGC4}
+
+The OGC events are shown in Figure~\ref{fig:ogc4_events}.
 
@@
     \begin{centering}
-        \includegraphics[width=0.8\columnwidth]{figures/ogc4_events.pdf}
+        \includegraphics[width=0.35\columnwidth]{figures/ogc4_events.pdf}
         \caption{
@@
 
-\section{$\pi(\mathcal{M}, D_{C})$ to $\pi(\mathcal{M}, z)$}
-
-See \cite{Hogg:1999:arXiv} for transforms
-
+The priors for the events are in $\pi(\Mc,V_c)$, however, we need $\pi(\Mc,z)$. To transform the priors, we use the following equation
 \begin{equation}
-\Omega_{\rm M}\equiv\frac{8\pi\,G\,\rho_0}{3\,H_0^2}
+    \pi(\Mc,z) = \pi(\Mc, V_c(z)) \frac{d V_c(z)}{dz}\, .
 \end{equation}
-\begin{equation}
-\Omega_{\Lambda}\equiv\frac{\Lambda\,c^2}{3\,H_0^2}
-\end{equation}
-\begin{equation}
-    D_H = c / H_0
-\end{equation}
-
-\begin{equation}
-\label{eq:ez}
-E(z)\equiv\sqrt{\Omega_{\rm M}\,(1+z)^3+\Omega_k\,(1+z)^2+\Omega_{\Lambda}}
-\end{equation}
-
-\begin{equation}
-D_{\rm C} = D_{\rm H}\,\int_0^z\frac{dz'}{E(z')}
-\end{equation}
-
-The comoving distance between two events at the same redshift or
-distance but separated on the sky by some angle $\delta\theta$ is
-$D_{\rm M}\,\delta\theta$ and the transverse comoving distance $D_{\rm
-M}$ (so-denoted for a reason explained below) is simply related to the
-line-of-sight comoving distance $D_{\rm C}$:
-\begin{equation}
-D_{\rm M} = \left\{
-\begin{array}{ll}
-D_{\rm H}\,\frac{1}{\sqrt{\Omega_k}}\,\sinh\left[\sqrt{\Omega_k}\,D_{\rm C}/D_{\rm H}\right] & {\rm for}~\Omega_k>0 \\
-D_{\rm C} & {\rm for}~\Omega_k=0 \\
-D_{\rm H}\,\frac{1}{\sqrt{|\Omega_k|}}\,\sin\left[\sqrt{|\Omega_k|}\,D_{\rm C}/D_{\rm H}\right] & {\rm for}~\Omega_k<0
-\end{array}
-\right.
-\end{equation}
-
-Angular diameter distance is related to the
-transverse comoving distance by
-\begin{equation}
-D_{\rm A} = \frac{D_{\rm M}}{1+z}
-\end{equation}
-
-The comoving volume element in solid
-angle $d\Omega$ and redshift interval $dz$ is
+We use the definition of $d V_c(z)/dz$ from \citep{Hogg:1999:arXiv}, given by 
 \begin{equation}
 dV_{\rm C}= D_{\rm H}\,\frac{(1+z)^2\,D_{\rm A}^2}{E(z)}\,d\Omega\,dz
-\end{equation}
-
- The integral of the comoving volume element
-from the present to redshift $z$ gives the total comoving volume,
-all-sky, out to redshift $z$
-
-\begin{equation}
-V_{\rm C} = \left\{
-\begin{array}{ll}
-  \left(\frac{4\pi\,D_{\rm H}^3}{2\,\Omega_k}\right)\,
-  \left[\frac{D_{\rm M}}{D_{\rm H}}\,
-  \sqrt{1+\Omega_k\,\frac{D_{\rm M}^2}{D_{\rm H}^2}}
-  -\frac{1}{\sqrt{|\Omega_k|}}\,
-  {\rm arcsinh}\left(\sqrt{|\Omega_k|}\,\frac{D_{\rm M}}{D_{\rm H}}\right)\right]
-  & {\rm for}~\Omega_k>0 \\
-  \frac{4\pi}{3}\,D_{\rm M}^3
-  & {\rm for}~\Omega_k=0 \\
-  \left(\frac{4\pi\,D_{\rm H}^3}{2\,\Omega_k}\right)\,
-  \left[\frac{D_{\rm M}}{D_{\rm H}}\,
-  \sqrt{1+\Omega_k\,\frac{D_{\rm M}^2}{D_{\rm H}^2}}
-  -\frac{1}{\sqrt{|\Omega_k|}}\,
-  {\rm arcsin}\left(\sqrt{|\Omega_k|}\,\frac{D_{\rm M}}{D_{\rm H}}\right)\right]
-  & {\rm for}~\Omega_k<0
-\end{array}
-\right.
 \end{equation}

### 2024-06-26 src/scripts/plot_mcz_grid.py
add mz plot

@@ -1,0 +1,101 @@
+from compas_python_utils.cosmic_integration.binned_cosmic_integrator.cosmological_model import CosmologicalModel
+from compas_python_utils.cosmic_integration.binned_cosmic_integrator.detection_matrix import DetectionMatrix
+from collections  import namedtuple
+import paths
+import os
+import numpy as np
+from typing import List
+import matplotlib.pyplot as plt
+from ogc4_interface.plotting.cmap_generator import CMAP
+from scipy.ndimage import gaussian_filter
+from data_cache import DataCache
+ROOT = "https://github.com/COMPAS-Surrogate/paper_data/raw/main/detection_matrix"
+
+
+DATA_URLS = [
+    f"{ROOT}/mczgrid_h5out_512M_aSF_0.0100_dSF_4.7000_mu_z_-0.2300_sigma_z_0.0000.h5",
+    f"{ROOT}/mczgrid_h5out_512M_aSF_0.0100_dSF_6.0000_mu_z_-0.2300_sigma_0_3.0000.h5",
+    f"{ROOT}/mczgrid_h5out_512M_aSF_0.0100_dSF_3.0000_mu_z_-0.0100_sigma_0_0.2500.h5"
+]
+
+
+MCZ_MODEL = namedtuple("mcz", ["matrix", "z", "sfr", "dPdlogZ"])
+
+
+redshift_range = [0, 10.0]
+logZ_range = [-12.0, 0] # metallicity
+mc_bins = np.linspace(3, 40, 50)
+redshift_bins = np.linspace(0, 1, 100)
+
+
+
+
+
+
+
+def load_data()->List[MCZ_MODEL]:
+    mcz_data = []
+    for url in DATA_URLS:
+        fname = DataCache(url).fpath
+        det_matrx = DetectionMatrix.from_h5(fname)
+        cosmo = CosmologicalModel(**det_matrx.cosmological_parameters)
+        mcz_data.append(MCZ_MODEL(det_matrx.rate_matrix, cosmo.redshift, cosmo.sfr, cosmo.dPdlogZ))
+    return mcz_data
+
+def plot_mcz_grid(
+        colors=["tab:green", "tab:orange", "tab:blue"],
+        cmaps=["Greens", "Oranges", "Blues"]
+):
+    data = load_data()
+    fig, axes = plt.subplots(1, 3, figsize=(15, 4))
+
+    axes[0].set_xlabel(r"$z$")
[…]

### 2024-06-26 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\subsection{MSSFR parameters}\label{sec:method_parameters}
-\avi{SECTION TAKEN FROM \citep{Riley:2023:ApJ}, FIGURE OUT OUT HOW CONDENSE}
-The parameters we chose to vary for this study are four free parameters for the calculation of the \ac{MSSFR} in the phenomenological model of \citet{Neijssel_2019}.
+\subsection{MSSFR parameters and $\Mc-z$ rates}\label{sec:method_parameters}
+Following \citep{Riley:2023:ApJ}, we vary four free parameters for the calculation of the \ac{MSSFR} in the phenomenological model of \citet{Neijssel_2019}.
 In this model, the \ac{MSSFR} is split into two parts, the star formation rate (SFR) and the metallicity distribution:
@@
 \bigskip\noindent
-We vary SFR parameters $a$ and $d$, and the metallicity distribution parameters $\alpha$ and $\sigma$, in this study, while fixing $b=2.77$, $c=2.9$, and $Z_0=0.035$.
-\resp{\citet{Neijssel_2019} demonstrated that a range of star-formation history models can be mimicked by varying only this subset of \ac{MSSFR} parameters.
-We therefore choose to vary the same parameters for consistency with their analysis.
-Moreover, reducing the number of parameters allows us to limit computational complexity for this proof-of-principle study.}
+We vary SFR parameters $a$ and $d$, and the metallicity distribution parameters $\alpha$ and $\sigma$. 
+We fix $b=2.77$, $c=2.9$, and $Z_0=0.035$.
 
-\subsection{Cosmic integration}
-Talk about how we run cosmic int,
-show plots of Mcz grid for different SF
+Varying these parameters in \COMPAS results in different merger rate distributions (see \citep{Riley:2023:ApJ, Neijssel_2019}).
+
+Plots of the SFR, metallicity distributions and resultant merger rate distributions for different $\Mc$ and $z$ are shown in Figure~\ref{fig:mcz_grid} for three different choices of parameters.
+
+\begin{figure*}[ht!]
+    \script{plot_mcz_grid.py}
+    \begin{centering}
+        \includegraphics[width=\textwidth]{figures/mcz_grid.pdf}
+        \caption{
+            McZ grid given differnt SFR parameters. Fiducial model is in X, 
+        }
+        \label{fig:mcz_grid}
+    \end{centering}
+\end{figure*}
 
@@
 
-\begin{equation}
+\begin{equation}\label{eq:weights}
     w_{i,b} = \frac{1}{K} \sum^K_{k=1} \frac{I(\{z^k_i,\Mc^k_i\} \in b)}{\pi(z_i^k,\Mc_i^k)} \ ,
@@
 
-One of the most popular BO algorithms is the Efficient Global Optimization (EGO) \cite{jones1998efficient}. 
-It uses GP as surrogate model and the Expected Improvement as the infill sampling criterion. This work is focused on BO algorithms using GP and its variants as surrogate models.
+One of the most popular BO algorithms is the Efficient Global Optimization (EGO). 
+It uses GP as surrogate model and the Expected Improvement as the infill sampling criterion.
 
@@
 
+We run an MCMC sampler using
 \begin{equation}
@@
 
+The KL divergence using the two \Lsurr should converge once the model is trained. 
+Additionally, the KL divergence using a reference posterior trainined with a lot of points should converge.
+This will indicate that the surrogate likelihood distribution has been adequately modelled in regions of parameter space.
+
 %--------------------------------------------------------
@@
 
-To validate our method, following Section 2.5 of \citep{}, we sample 100 $\lambda$ from a $\pi(\lambda)$, and generate $\vec{\mathcal{D}}$, unique mock \COMPAS populations for each $\lambda$.
+To validate our method, we conduct the the study described in Section 2.5 of \citep{Riley:2023:ApJ}.
+
+We simulate the evolution of 32 million binary systems using \COMPAS (settings specified in Appendix~\ref{sec:appendix_COMPAS_fiducial}), resulting in a population of approximately 68 thousand DCOs that merge within the age of the Universe, $\sim13.8\unit{Gyr}$.
+We sample 100 $\lambda$ from a $\pi(\lambda)$, and generate $\vec{\mathcal{D}}$, unique mock \COMPAS populations for each $\lambda$.
 We build surrogate likelihoods for each dataset, and use them to approximate $p(\lambda|\mathcal{D})$.
@@
 
-Finally, Figure~\ref{} shows the JS diveregence
-
-The catalogue of LIGO, Virgo, and KAGRA (LVK) gravitational wave observations (\citet{GWTC-2_1_zenodo}, \citet{GWTC-3_zenodo}) provides us with a number of merger detections from a population of \acp{DCO}.
-For each LVK event we know, among other things, the chirp mass of the merging binary and the merger redshift, up to a measurement uncertainty.
-We only include confident detections with a minimum astrophysical probability $p_{astro} \ge 0.95$; although less confident events could be included \citep[e.g.][]{Farr_2015}, they will typically contribute little information due to greater measurement uncertainties, so we omit them for this proof-of-concept study.
-
-Once we validate our method, we can use our surrogate model and the LVK data to infer real physical constraints on the parameters under study, with the important caveat that we have fixed the astrophysical parameters, and any errors in those could bias the inferred \ac{MSSFR} model values.
-
-Include corners of variable LnL + normal LnL
-Include KL divergence (low number points --> high points)
-Include PP plot
+Finally, Figure~\ref{fig:kl_distances} shows the JS diveregence
 
@@
 \subsection{LVK analysis}
-\paragraph{}{Data}
+\paragraph{Data}
 
+The catalogue of LIGO, Virgo, and KAGRA (LVK) gravitational wave observations (\citet{GWTC-2_1_zenodo}, \citet{GWTC-3_zenodo}) provides us with a number of merger detections from a population of \acp{DCO}.
+ For each LVK event we know, among other things, the chirp mass of the merging binary and the merger redshift, up to a measurement uncertainty.
 We get the data from the \ogc.
-Plots of the event $\Mc,z$ median estimates and $X-\sigma$ uncertainty can be found in Figure~\ref{fig:ogc4_events} in the Appendix~\ref{}. 
+Plots of the event $\Mc,z$ median estimates and $X-\sigma$ uncertainty can be found in Figure~\ref{fig:ogc4_events}. 
 Filtering the events based on the OGC definition of $p(bbh)\geq0.95$ and ensuring that the median value of the $p(\Mc,z|d_i)$ lies within $\Mc\in[4,45]$, $z\in[0,1]$, we get XX events. 
-These are colored green in Fig~\ref{fig:ogc4_events}.
+Less confident events could be included \citep[e.g.][]{Farr_2015}, they will typically contribute little information due to greater measurement uncertainties.
+These events that we use in our analysis are colored green in Fig~\ref{fig:ogc4_events}.
 
-Using Equation~\ref{}, we obtain the weights for each posterior. 
+\begin{figure}[ht!]
+    \script{plot_ogc4_events.py}
+    \begin{centering}
+        \includegraphics[width=0.8\columnwidth]{figures/ogc4_events.pdf}
+        \caption{
+            The events and uncertainty as predicted by Ogc4
+        }
+        \label{fig:ogc4_events}
+    \end{centering}
+\end{figure}
+
+The priors for the events are in $\pi(\Mc,V_c)$, however, we need $\pi(\Mc,z)$. To transform the priors, we use the following equation
+\begin{equation}
+    \pi(\Mc,z) = \pi(\Mc, V_c(z)) \frac{d V_c(z)}{dz}\, .
+\end{equation}
+We use the definition of $d V_c(z)/dz$ from \citep{Hogg:1999:arXiv}, given by 
+\begin{equation}
+dV_{\rm C}= D_{\rm H}\,\frac{(1+z)^2\,D_{\rm A}^2}{E(z)}\,d\Omega\,dz
+\end{equation}
+
+Using Equation~\ref{eq:weights}, we obtain the weights for each posterior. 
 A plot of each event's weights summed together is shown in Figure~\ref{fig:ogc4_weights}.
@@
 
-Describe OGC4 data
-
-\subsection{Results}
-Describe LVK results, compare with Riley et al.
+\paragraph{Posteriors}
+pass
 
@@
 \appendix
-
-\section{Interfacing with OGC4}
-
-The OGC events are shown in Figure~\ref{fig:ogc4_events}.
-
-\begin{figure}[ht!]
-    \script{plot_ogc4_events.py}
-    \begin{centering}
-        \includegraphics[width=0.35\columnwidth]{figures/ogc4_events.pdf}
-        \caption{
-            The events and uncertainty as predicted by Ogc4
-        }
-        \label{fig:ogc4_events}
-    \end{centering}
-\end{figure}
-
-The priors for the events are in $\pi(\Mc,V_c)$, however, we need $\pi(\Mc,z)$. To transform the priors, we use the following equation
-\begin{equation}
-    \pi(\Mc,z) = \pi(\Mc, V_c(z)) \frac{d V_c(z)}{dz}\, .
-\end{equation}
-We use the definition of $d V_c(z)/dz$ from \citep{Hogg:1999:arXiv}, given by 
-\begin{equation}
-dV_{\rm C}= D_{\rm H}\,\frac{(1+z)^2\,D_{\rm A}^2}{E(z)}\,d\Omega\,dz
-\end{equation}
 
@@
 
+\section{Sampler settings}\label{sec:appendix_sampler_settubgs}
+Here we describe the sampler settings we use. 
+We also plot some trace plots.
+
 \end{document}

### 2024-06-27 src/scripts/plot_simulation_posterior.py
pushing changes

@@ -64,7 +64,7 @@ class PosteriorDatasets:
     def from_npz(cls, fname):
         data = np.load(fname)
         posteriors = [p for p in data["posteriors"]]
-        return cls(posteriors, data["truths"], data["params"], data["labels"])
+        return cls(posteriors, data["reference_param"], data["params"], data["labels"])
 
     @classmethod
     def from_cache(cls):
@@ -74,7 +74,7 @@ class PosteriorDatasets:
         return {
             "chains": self.chains[0:2],
             "chainLabels": self.chainLabels[0:2],
-            "truths": self.truths,
+            "reference_param": self.truths,
             "paramNames": self.paramNames,
             "paramRanges": self.paramRanges,
             "reference": self.chains[2],

### 2024-06-27 src/scripts/pygtc.py

@@ -30,7 +30,7 @@ REF_COLOR = 'lightgray'
 def plotGTC(chains, **kwargs):
     r"""Make a great looking Giant Triangle Confusogram (GTC) with one line of
     code! A GTC is like a triangle (or corner) plot, but you get to put as
-    many sets of data, and overlay as many truths as you like. That's what can
+    many sets of data, and overlay as many reference_param as you like. That's what can
     make it so *confusing*!
 
     Parameters
@@ -66,23 +66,23 @@ def plotGTC(chains, **kwargs):
     truths : list-like[nDims] or [[nDims], ...]
         A list of parameter values, one for each parameter in `chains` to
         highlight in the GTC parameter space, or a list of lists of values to
-        highlight in the parameter space. For each set of truths passed to
-        `truths`, there must be a value corresponding to every dimension in
+        highlight in the parameter space. For each set of reference_param passed to
+        `reference_param`, there must be a value corresponding to every dimension in
         `chains`, although any value may be ``None``. Default is ``None``.
 
     truthLabels : list-like[nTruths]
-        A list of labels, one for each list passed to truths. truthLabels
+        A list of labels, one for each list passed to reference_param. truthLabels
         supports LaTex commands enclosed in $..$. Additionally, you can pass
         ``None`` as a label. Default is ``None``.
 
     truthColors : list-like[nTruths]
-        User-defined colors for the truth lines, must be one per set of truths
-        passed to `truths`. Default color is gray ``#4d4d4d`` for up to three
+        User-defined colors for the reference_param lines, must be one per set of reference_param
+        passed to `reference_param`. Default color is gray ``#4d4d4d`` for up to three
         lines.
 
     truthLineStyles : list-like[nTruths]
-        User-defined line styles for the truth lines, must be one per set of
-        truths passed to `truths`. Default line styles are
+        User-defined line styles for the reference_param lines, must be one per set of
+        reference_param passed to `reference_param`. Default line styles are
         ``['--', ':', 'dashdot']``.
 
     priors : list of tuples [(mu1, sigma1), ...]
@@ -132,7 +132,7 @@ def plotGTC(chains, **kwargs):
     legendMarker : string
         Options are ``'All'``, ``'None'``, ``'Auto'``. ``'All'`` and ``'None'``
         force-show or force-hide all label markers. ``'Auto'`` shows label
-        markers if two or more truths are plotted.
[…]

### 2024-11-12 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{showyourwork}
+\usepackage{tikz}
+\usepackage{float}
+\usepackage{algorithm}
+% \usepackage{algorithmic}
+\usepackage{algpseudocode}
 
-\usepackage{float}
-\makeatletter
-\let\newfloat\newfloat@ltx
-\makeatother
-
-\usepackage{algorithm}
-\usepackage{algorithmic}
-
-\usepackage{tikz}
-\usetikzlibrary{shapes.geometric, arrows}
-
-\setlength{\marginparwidth}{1.5cm}
-
-% table column types
-\usepackage{array}
-\newcolumntype{E}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{F}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{G}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-% abbreviations/acronyms
-\acrodef{ANN}{artificial neural network}
-\acrodef{DCO}{double compact object}
-\acrodef{MCMC}{Markov Chain Monte Carlo}
-\acrodef{MSSFR}{metallicity-specific star formation rate}
-\acrodef{LVK}{LIGO--VIRGO--KAGRA}
-
-\newcommand\COMPAS{{\sc{COMPAS}}\xspace}
-\newcommand*\diff{\mathop{}\!\mathrm{d}}
-\newcommand{\MSFR}{\ensuremath{{M}_{\rm{SFR}}}\xspace}
-\newcommand{\ts}{\ensuremath{{t}_{\rm{s}}}\xspace}
-\newcommand{\tsup}{\textsuperscript}
-\newcommand{\Vc}{\ensuremath{{V}_{\rm{c}}}\xspace}
-
-\newcommand{\monei}{\ensuremath{m_{1,\rm{i}}}\xspace}
-\newcommand{\mtwoi}{\ensuremath{m_{2,\rm{i}}}\xspace}
-\newcommand{\monef}{\ensuremath{m_{1,\rm{f}}}\xspace}
-\newcommand{\mtwof}{\ensuremath{m_{2,\rm{f}}}\xspace}
-\newcommand{\ai}{\ensuremath{a_{\rm{i}}}\xspace}
-\newcommand{\qi}{\ensuremath{q_{\rm{i}}}\xspace}
-\newcommand{\Zi}{\ensuremath{Z_{\rm{i}}}\xspace}
-\newcommand{\vk}{\ensuremath{v_{\rm{k}}}\xspace}
-\newcommand{\thetak}{\ensuremath{{\theta}_{\rm{k}}}\xspace}
-\newcommand{\phik}{\ensuremath{{\phi}_{\rm{k}}}\xspace}
-\newcommand{\ei}{\ensuremath{{e}_{\rm{i}}}\xspace}
-
-\newcommand{\Msun}{\ensuremath{\,\rm{M}_{\odot}}\xspace}
-\newcommand{\Zsun}{\ensuremath{\,\rm{Z}_{\odot}}\xspace}
-\newcommand{\kms}{\ensuremath{\,\rm{km}\,\rm{s}^{-1}}\xspace}
-\newcommand{\AU}{\ensuremath{\,\mathrm{AU}}\xspace}
-\newcommand{\Mc}{\ensuremath{\mathcal{M}}\xspace}
-
-\newcommand{\Lmain}{\ensuremath{\mathcal{L}(\mathcal{D}|\lambda)}\xspace}
-\newcommand{\Lp}{\ensuremath{\mathcal{L}_{\rm p}}\xspace}
-\newcommand{\Le}{\ensuremath{\mathcal{L}_{\rm event}}\xspace}
-\newcommand{\pg}{\ensuremath{p_{\rm grid}}\xspace}
-\newcommand{\Lsurr}{\ensuremath{\mathcal{L}^{\star}(\lambda)}\xspace}
-
-%%%%%%%%%%%%%%% SYMBOLS
-
-% Projects:
-\newcommand{\project}[1]{\textsf{#1}}
-
-\newcommand{\python}{\project{Python}}
-\newcommand{\jupyter}{\project{Jupyter}}
-
-\newcommand{\exoplanet}{\project{exoplanet}}
-\newcommand{\lightkurve}{\project{lightkurve}}
-\newcommand{\starry}{\project{starry}}
-\newcommand{\pymc}{\project{PyMC3}}
-\newcommand{\pymcextra}{\project{pymc3-ext}}
-\newcommand{\celerite}{\project{celerite}}
-\newcommand{\astropy}{\project{astropy}}
-\newcommand{\scipy}{\project{scipy}}
-\newcommand{\jupytext}{\project{jupytext}}
-\newcommand{\sphinx}{\project{sphinx}}
-\newcommand{\jupyterbook}{\project{Jupyter-book}}
-\newcommand{\arviz}{\project{ArviZ}}
-\newcommand{\nbconvert}{\project{nbconvert}}
-\newcommand{\numpy}{\project{numpy}}
-\newcommand{\pandas}{\project{pandas}}
-\newcommand{\matplotlib}{\project{matplotlib}}
-\newcommand{\corner}{\project{corner}}
-\newcommand{\emcee}{\project{emcee}}
-\newcommand{\trieste}{\project{trieste}}
-\newcommand{\tensorflow}{\project{TensorFlow}}
-\newcommand{\pycbc}{\project{PyCBC}}
-
-\newcommand{\LVK}{\project{LVK}}
-\newcommand{\mast}{\project{MAST}}
-\newcommand{\exofop}{\project{ExoFOP}}
-\newcommand{\tessAtlas}{\project{TESS Atlas}}
-\newcommand{\ogc}{\project{4-OGC}}
-
-% math
-\newcommand{\T}{\ensuremath{\mathrm{T}}}
-\newcommand{\dd}{\ensuremath{ \mathrm{d}}}
-% \newcommand{\unit}[1]{{\ensuremath{ \mathrm{#1}}}}
-\newcommand{\bvec}[1]{{\ensuremath{\boldsymbol{#1}}}}
-
-\DeclareMathOperator{\invG}{Inv-\mathnormal{\Gamma}}
-\DeclareMathOperator{\N}{\mathcal{N}}
-\DeclareMathOperator{\U}{\mathcal{U}}
-\DeclareMathOperator{\Un}{\mathcal{U}}
-\DeclareMathOperator{\Par}{\mathcal{P}ar}
-\DeclareMathOperator{\tmin}{\mathnormal{t_{\rm min}}}
-\DeclareMathOperator{\tmax}{\mathnormal{t_{\rm max}}}
-
-%% affiliation shortcuts
-\newcommand{\SPA}{School of Physics and Astronomy, Monash University, Clayton VIC 3800, Australia}
-\newcommand{\OzGravMonash}{OzGrav: The ARC Centre of Excellence for Gravitational Wave Discovery, Clayton VIC 3800, Australia}
-\newcommand{\AMNH}{Department of Astrophysics, American Museum of Natural History, New York, NY 10024, USA}
-\newcommand{\CCA}{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}
-\newcommand{\CUNY}{Graduate Center, City University of New York, 365 5th Avenue, New York, NY 10016, USA}
-\newcommand{\BMCC}{Department of Science, BMCC, City University of New York, New York, NY 10007, USA}
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%
-
-\definecolor{orange-red}{rgb}{1.0, 0.27, 0.0}
-\newcommand{\jeff}[1]{\textbf{\textcolor{orange-red}{#1}}}
-\newcommand{\ilya}[1]{\textbf{\textcolor{magenta}{#1}}}
-\newcommand{\avi}[1]{\textbf{\textcolor{cyan}{#1}}}
-\newcommand{\chayan}[1]{\textbf{\textcolor{green}{#1}}}
-\newcommand{\resp}[1]{#1}
+\input{my-style}
 
@@
 
-% Authors
-\author{Avi Vajpeyi}
-\affiliation{UoA}
-\affiliation{NZ Gravity}
-\affiliation{ARC Centre of Excellence for Gravitational Wave Discovery -- OzGrav, Australia}
-
-\author{Jeff Riley}
-\affiliation{School of Physics and Astronomy, Monash University, Clayton, Victoria 3800, Australia}
-\affiliation{ARC Centre of Excellence for Gravitational Wave Discovery -- OzGrav, Australia}
-\author{Ilya Mandel}
-\affiliation{School of Physics and Astronomy, Monash University, Clayton, Victoria 3800, Australia}
-\affiliation{ARC Centre of Excellence for Gravitational Wave Discovery -- OzGrav, Australia}
-\author{Chayan Chatterjee}
-\affiliation{}
-
+\input{authors}
 \correspondingauthor{Avi Vajpeyi}
@@
 \label{sec:intro}
-Over the past decade, improvements in detector sensitivity have led to an exponential increase in the rate of GW detections.
-With next-generation detectors on the horizon (such as CE, ET, and LISA), the number of GW detections will increase even more.
-The most recent catalogue of GW contains
-XX events, the majority of them being Binary Black Holes
-(BBHs), with XX public alerts for GW that have yet to be confirmed as astrophysiocal events.
-As more GW events are observed, more features
-in the distribution of GW events await to be discovered.
-One of the most common ways to understand the observed GW population is through phenomenological modelling.
-However, as the size of the catalogues increases, so does the catalogue complexity, making it challenging to develop interpretable, parameterisable phenomenological models, flexible enough to explain the features in the catalogue.
 
-Alternatively, we can consider using forward models to simulate catalogues of GW events and compare the simulations to observed catalogues.
-The forward modelling approach would provide astronomers with constraints on easy-to-interpret physical parameters that describe the initial conditions of the population of merging binaries, such as distributions describing the initial stellar mass, metallicity, and binary separation.
-These parameters also govern simulated physical processes like wind mass loss, supernova characteristics, and binary interactions such as mass transfer and common envelope efficiency.
-However, the generation of synthetic catalogues from the various rapid population sysnthesis software packages can be expensive, making brute force methods that explore all possible parameter values expensive (for example, see  Broekgaarden et al.
-(2022)).
-Another alternative, as proposed by Riley et al. (2023) involves building surrogate tools to bridge the computational complexities of simulating a state space large enough
-to allow us to infer astrophysical constraints from the observed catalogs.
+Over the past decade, improvements in detector sensitivity have led to a rapid increase in the rate of gravitational wave (GW) detections~\citep{GWTC1, GWTC2, GWTC3}. 
+With next-generation detectors on the horizon, such as Cosmic Explorer (CE), Einstein Telescope (ET), and the Laser Interferometer Space Antenna (LISA), the number of GW detections is expected to increase even more dramatically~\citep{ET_design_report, ET_science_case, CE_horizon_study, LISA_science_case, LISA_red_book}.
+The most recent catalog of GW events, GWTC-3, contains 90 confident detections, the majority being binary black holes (BBHs), with additional public alerts for GW events yet to be confirmed as astrophysical~\citep{GWTC3, GraceDB2024}.
+As more GW events are observed, new features in the distribution of these events await discovery, offering unprecedented insights into astrophysical populations and processes.
+A method to infer properties about the populations of GW events, while accouting for selection bias and uncertain source parameters, is using hierarchical Bayesian inference~\citep{Mandel:2019:MNRAS, Vitale:2022:hgwa}.
 
-Riley et al. (2023) generated a small number of ``training" states using a
-targeted selection of initial conditions and evolutionary
-parameters.
-The training states were then used to teach an interpolant how to
-map initial conditions and evolutionary parameters to
-the final state, thus avoiding the computational time
-of calculating the intermediate steps.
+In the inference process, a common approach to understanding the observed GW population is through phenomenological modeling~\citep{Talbot:2017:PhRvD, Talbot:2018:ApJ, gwpopulation, Wysocki:2019:PhRvD, Fishbach:2018:ApJL, GWTC2PopInf, GWTC3PopInf, Mastrogiovanni:2023:arXiv}.
+However, as the size and complexity of GW catalogs increase, developing interpretable and parameterizable phenomenological models that are flexible enough to explain the emerging features becomes increasingly challenging~\citep{Golomb:2023:PhRvD}.
 
-In their work, they explored four parameters governing the cosmic metallicity-specific star formation rate (MSSFR) in the model of Neijssel et al. (2019).
+Some alternative approaches involve using non-parameteric models~\citep{Li:2021:ApJ, Rinaldi:2024:A&A, Heinzel:2024:arXiv, Callister:2024:PhRvX}, and forward modelling to simulate catalogs of GW events and compare these simulated catalogs to observed catalogs~\citep{Wong:2023:ApJ, Riley:2023:ApJ}.
+This forward modeling approach provides astronomers with constraints on interpretable physical parameters that describe the initial conditions of progenitor stellar populations. 
+These parameters include distributions of initial stellar mass, metallicity, and binary separation, as well as those governing simulated physical processes such as wind mass loss, supernova characteristics, and binary interactions such as mass transfer and common envelope efficiency.
+However, generating synthetic catalogs from various rapid population synthesis software packages, e.g., \COMPAS~\citep{Compas:2022:JOSS}, \BPASS~\citep{BPASS:2017:PASA}, \COSMIC~\citep{COSMIC:2020:ApJ}, \STARTRACK~\citep{StarTrack:2008:ApJS}, and \POSYDEN~\citep{POSYDON:2023:ApJS}, can be computationally expensive, making brute-force methods that explore all possible parameter values impractical.
+To address this challenge, \citet{Riley:2023:ApJ} proposed building surrogate tools to bridge the computational complexities of simulating a state space large enough to allow inference of astrophysical constraints from observed catalogs.
 
-We build upon \citep{Riley:2023:ApJ}'s work, and
+\citet{Riley:2023:ApJ} generated a set of ``training'' merging binary populations by sampling from a four parameters that govern the metallicity-specific star formation rate (MSSFR), \MSSFR, in the model proposed by \citet{Neijssel_2019}.
+These populations were used to construct neural network interpolants that map MSSFR parameters to merger rates, thereby circumventing the computational burden of evolving the binaries from initial conditions.
+The interpolants were then employed within a Markov chain Monte Carlo (MCMC) framework to approximate the posterior distribution of MSSFR parameters, using both simulated data and observations from GWTC-3.
 
-The remainder of this paper is organised as follows.
-Section 2 presents a description of the tool we construct
-for this proof-of-concept study, and the method used
-to train the tool.
-We present and discuss our results in Section 3.
-We provide some concluding remarks in  Section 4
+Our study extends the work of \citet{Riley:2023:ApJ} by implementing Gaussian processes (GPs) to map MSSFR parameters with an observed dataset to a likelihood function. 
+This likelihood is then utilized in an MCMC to approximate the MSSFR parameter posteriors given the observed dataset. 
+Crucially, we aim to enhance the efficiency of \citet{Riley:2023:ApJ}'s approach by reducing the size of the training set required for the GP. 
+To achieve this, we use an ``active-learning'' approach to train the GP, where we iteratively employ Bayesian hyperoptimization for acquiring MSSFR parameters, based on the GP's performance.
+This active learning approach is expected to be particularly advantageous in more complex scenarios, where generating each population is computationally expensive, potentially offering a more scalable and resource-efficient method for exploring high-dimensional parameter spaces in astrophysical population synthesis.
+
+The remainder of this paper is organized as follows. 
+Section~\ref{sec:methods} presents a description of the active learning methodology we construct for this proof-of-concept study along with the Bayesian framework used. 
+We apply our method to simulated datasets and the GWTC-3 in Section~\ref{sec:results}. Finally, we provide concluding remarks and discuss future directions in Section~\ref{sec:discussion}.
 
 %--------------------------------------------------------
-\section{Method}
+\section{Method}\label{sec:methods}
 
-\subsection{MSSFR parameters and $\Mc-z$ rates}\label{sec:method_parameters}
-Following \citep{Riley:2023:ApJ}, we vary four free parameters for the calculation of the \ac{MSSFR} in the phenomenological model of \citet{Neijssel_2019}.
-In this model, the \ac{MSSFR} is split into two parts, the star formation rate (SFR) and the metallicity distribution:
+\subsection{Forward model}\label{sec:method_parameters}
+Following \citet{Riley:2023:ApJ}, we use \COMPAS to simulate binaries populations and vary four free parameters for the calculation of the metallicity-specific star formation rate (MSSFR) in the phenomenological model of \citet{Neijssel_2019}.
+In this model, the MSSFR is split into two parts, the star formation rate (SFR) and the metallicity distribution.
+
+We assume the analytical star formation rate, proposed by \citet{Madau_2014}, 
+\begin{equation}
+  \frac{\diff^2 M_{\rm SFR}}{\diff t\,dV_c}(z | a,b,c,d) =
+  \frac{a\ (1 + z)^b}{1 + \big(\frac{1 + z}{c}\big)^d}\,\bigg[ \frac{\mathrm{M}_\odot}{\mathrm{yr}\ \mathrm{Mpc}^{3}}\bigg]
+  \ ,
+\end{equation}
 
@@
 where the SFR is given by:
-
-\begin{equation}
-  \frac{\diff^2 \MSFR}{\diff \ts \diff \Vc} =
-  a\frac{(1 + z)^b}{1 + (\frac{(1 + z)}{c})^d}\,
-  \mathrm{M}_\odot\ \mathrm{yr}^{-1}\ \mathrm{Mpc}^{-3},
-\end{equation}
 
@@
 
-Varying these parameters in \COMPAS results in different merger rate distributions (see \citep{Riley:2023:ApJ, Neijssel_2019}).
+Varying these parameters in \COMPAS results in different merger rate distributions~\citep{Riley:2023:ApJ, Neijssel_2019}.
 
@@
 \subsection{Bayesian Model}
+
+% The likelihood of obtaining \Nobs observations from a population of events over on observing time \Tobs, each
+% described by redshift $z$ and chirp-mass \Mc, in a data collection  $\mathcal{D} \equiv \{D_1, D_2, ... D_{N_{\mathrm{obs}}}\}$, assuming a
+% constant rate  selection biases is given by (see
+% Mandel et al. (2019); Vitale et al. (2022)
+
 We use the Bayesian model described in \citep{Riley:2023:ApJ}.
-Given $N_{\mathrm{obs}}$ observations from data  $\mathcal{D} \equiv \{D_1, D_2, ...
-D_{N_{\mathrm{obs}}}\}$, we define the likelihood $\mathcal{L}$ that we will see data $\mathcal{D}$ for a particular set of parameters $\lambda$ as
+Given $\Nobs$ observations from a population, with each of the $D_i$ observations described by redshift $z_i$ and chirp-mass $\Mc_i$, making a data collection $\mathcal{D} \equiv \{D_1, D_2, ...
+D_{N_{\mathrm{obs}}}\}$, we define the likelihood $\mathcal{L}$ to observe data $\mathcal{D}$ for a particular set of parameters $\lambda$ as
 \begin{equation}
@@
 
-\bigskip
 The first term of Equation~\ref{eq:methods_likelihood} is a Poisson likelihood on there being $N_\mathrm{obs}$ detections,
@@
 \begin{equation}
-  %\notag
-  \log{ \Lp\left(N_{\mathrm{obs}}|\lambda\right)} = - \mu(\lambda) + N_{\mathrm{obs}}\log\mu(\lambda) ,
+  \log{ \Lp\left(N_{\mathrm{obs}}|\lambda\right)} =  N_{\mathrm{obs}}\log\mu(\lambda)- \mu(\lambda)\, ,
   \label{eq:methods_likelihood_term1}
@@
 
-\bigskip\noindent
 where $\mu(\lambda)$ is the expected number of observations and we omit terms that depend on the data only and therefore disappear on normalisation, such as $\log(N_{\mathrm{obs}}!)$ and permutation coefficients.
@@
 \begin{equation}
-  %\notag
   \Le(D_i|\lambda) = \pg(z=z_i,\mathcal{M}=\mathcal{M}_{i}|\lambda),
@@
 \bigskip\noindent
-where $\pg(z=z_i,\mathcal{M}=\mathcal{M}_{i}|\lambda)$ is the entry in the probability distribution matrix $\pg(z,\mathcal{M}|\lambda)$ in the $z$ and $\mathcal{M}$ bin of the observed $i^{th}$ event.
+where $\pg(z=z_i,\mathcal{M}=\mathcal{M}_{i}|\lambda)$ is the entry in the probability distribution matrix $\pg(z,\mathcal{M}|\lambda)$ in the $z$ and $\mathcal{M}$ bin of the observed $i^{th}$ event (for e.g., one pixel from the Figure~\ref{fig:mcz_grid}c).
 
@@
 where $I(\{z^k_i,\Mc^k_i\} \in b$ is an indicator function that evaluates to 1 if sample $\{z^k_i,\Mc^k_i\}$ falls into bin $b$ and $0$ otherwise.
+\avi{Normalise these weights}
 Then the probability $\Le(D_i|\lambda)$ can be evaluated as a sum over bins,
@@
 In the next section, we discuss a surrogate model for \Lmain that is cheaper to sample with.
+
+\avi{talk about selection effects}
 
@@
 
-\begin{algorithm}
-\caption{Basic pseudo-code for Bayesian optimization}
-\label{alg:BayesOpt}
-\begin{algorithmic}[1] % [1] turns on line numbering
-    \STATE Place Gaussian process $\mathcal{GP}$ prior on $\Lmain$
-    \STATE Observe $\Lmain$ at $N_0$ points randomly sampled.
-    \WHILE{$n \le N$}
-        \STATE Update the posterior probability distribution on $\Lmain$ using all available data
-        \STATE Let $\lambda_{\rm best}$ be a maximizer of the acquisition function over $\lambda$, where the acquisition function is computed using the current posterior distribution.
-        \STATE Observe $L_{\rm best} = \mathcal{L}(\mathcal{D}|\lambda_{\rm best})$.
-        \STATE Increment $n$
-    \ENDWHILE
-    \STATE Return a solution: $\Lsurr \sim \mathcal{GP}(\lambda)$
-\end{algorithmic}
-\end{algorithm}
+\input{algo}
 
@@
 
-\tikzstyle{startstop} = [rectangle, rounded corners, 
-minimum width=3cm, 
-minimum height=1cm,
-text centered, 
-draw=black, 
-fill=red!30]
-
-\tikzstyle{process} = [rectangle, 
-minimum width=3cm, 
-minimum height=1cm, 
-text centered, 
-text width=3cm, 
-draw=black, 
-fill=orange!30]
-
-\tikzstyle{decision} = [diamond, 
-minimum width=3cm, 
-minimum height=1cm, 
-text centered, 
-draw=black, 
-fill=green!30]
-\tikzstyle{arrow} = [thick,->,>=stealth]
-
-\begin{tikzpicture}[node distance=2cm]
-
-\node (start) [startstop] {Sample $\pi(\lambda)$};
-\node (compute) [process, below of=start] {Compute $\Lmain$};
-\node (train) [process, below of=compute] {Train \Lsurr};
-\node (sample) [process, below of=train] {Sample 
-$p(\lambda|\mathcal{D})\sim\Lsurr\pi(\lambda)$};
-\node (check) [decision, below of=sample, yshift=-0.5cm] {$\mathcal{O}(\lambda) > N?$};
-
-\node (stop) [startstop, below of=check] {Stop};
-\node (acquire) [process, right of=check, xshift=2cm] {Acquire $\lambda$};
-
-\draw [arrow] (start) -- (compute);
-\draw [arrow] (compute) -- (train);
-\draw [arrow] (train) -- (sample);
-\draw [arrow] (sample) -- (check);
-\draw [arrow] (check) -- node[anchor=east] {no} (acquire);
-\draw [arrow] (check) -- node[anchor=south] {yes} (stop);
-\draw [arrow] (acquire) |- (compute);
-
-\end{tikzpicture}
+\input{flowchart}
 
@@
 
-where $u^i = \frac{\mu^i - \mu_{\rm{best}}}{\sigma^i}$ and $\mu_{\rm{best}}$ is the best value of $\mu$ so far.
+where $u^i = {(\mu^i - \mu{\rm{best})}}/{\sigma^i}$ and $\mu_{\rm{best}}$ is the best value of $\mu$ so far.
 
@@
 \section{Inference with surrogate model}
-
+\label{sec:results}
 \subsection{Simulation study}
@@
 \section{Discussion}
+\label{sec:discussion}
 
@@
 Thankfully, as OzSTAR is powered by wind energy from Iberdrola Australia; the electricity for computations produces negligible carbon waste.
-A.V. is supported by the Australian Research Council (ARC) Centre of Excellence CE170100004.
-
 4-OGC.
@@
 
+\section{Posteriors}
+
+% \begin{figure}[htbp]
+%     \centering
+%     \begin{subfigure}{0.45\textwidth}
+%         \centering
+%         \includegraphics[width=\textwidth]{figures/5M_50_940pts.pdf}
+%         \caption{First image}
+%         \label{fig:image1}
+%     \end{subfigure}
+%     \hfill
+%     \begin{subfigure}{0.45\textwidth}
+%         \centering
+%         \includegraphics[width=\textwidth]{figures/32M_50_780pts.pdf}
+%         \caption{Second image}
+%         \label{fig:image2}
+%     \end{subfigure}
+%     \caption{Posteriors for MSSFR parameters using LVK data}
+%     \label{fig:two_images}
+% \end{figure}
+
+\begin{figure}[htb]
+    \centering
+    \begin{minipage}{0.45\columnwidth}
+        \includegraphics[width=\linewidth]{figures/5M_50_940pts.pdf}
+        \caption{Caption for subfigure (a)}
+    \end{minipage}\hfill
+    \begin{minipage}{0.45\columnwidth}
+        \includegraphics[width=\linewidth]{figures/32M_50_780pts.pdf}
+        \caption{Caption for subfigure (b)}
+    \end{minipage}
+    \caption{Main caption for the entire figure}
+    \label{fig:main}
+\end{figure}
+
 \end{document}

### 2024-11-12 src/tex/algo.tex
[showyourwork] overleaf sync

\begin{algorithm}
\caption{Basic pseudo-code for Bayesian optimization}
\label{alg:BayesOpt}
\begin{algorithmic}[1] % [1] turns on line numbering
    \State Place Gaussian process $\mathcal{GP}$ prior on $\Lmain$
    \State Observe $\Lmain$ at $N_0$ points randomly sampled.
    \While{$n \leq N$}
        \State Update the posterior probability distribution on $\Lmain$ using all available data
        \State Let $\lambda_{\text{best}}$ be a maximizer of the acquisition function over $\lambda$, where the acquisition function is computed using the current posterior distribution.
        \State Observe $L_{\text{best}} = \mathcal{L}(\mathcal{D}|\lambda_{\text{best}})$.
        \State Increment $n$
    \EndWhile
    \State Return a solution: $\Lsurr \sim \mathcal{GP}(\lambda)$
\end{algorithmic}
\end{algorithm}

### 2024-11-12 src/tex/flowchart.tex
[showyourwork] overleaf sync

\begin{figure}
    \centering
    \begin{tikzpicture}[node distance=0.7cm and 1.5cm]
        \node (start) [startstop] {Sample $\pi(\lambda)$};
        \node (compute) [process, below=of start] {Compute $\Lmain$};
        \node (train) [process, below=of compute] {Train $\Lsurr$};
        \node (sample) [process, below=of train] {Sample\\$p(\lambda|\mathcal{D})\sim\Lsurr\pi(\lambda)$};
        \node (check) [decision, below=of sample] {$\mathcal{O}(\lambda) > N?$};
        \node (stop) [startstop, below=of check] {Stop};
        \node (acquire) [process, right=of check] {Acquire $\lambda$};

        \draw [arrow] (start) -- (compute);
        \draw [arrow] (compute) -- (train);
        \draw [arrow] (train) -- (sample);
        \draw [arrow] (sample) -- (check);
        \draw [arrow] (check) -- node[above] {no} (acquire);
        \draw [arrow] (check) -- node[left] {yes} (stop);
        \draw [arrow] (acquire) -| (compute);
    \end{tikzpicture}
    \caption{Flowchart of the process}
    \label{fig:flowchart}
\end{figure}

### 2024-11-12 src/tex/algo.tex
[showyourwork] overleaf sync

@@
         \State Update the posterior probability distribution on $\Lmain$ using all available data
-        \State Let $\lambda_{\text{best}}$ be a maximizer of the acquisition function over $\lambda$, where the acquisition function is computed using the current posterior distribution.
-        \State Observe $L_{\text{best}} = \mathcal{L}(\mathcal{D}|\lambda_{\text{best}})$.
+        \State Let $\lambda_{\textrm{best}}$ be a maximizer of the acquisition function over $\lambda$, where the acquisition function is computed using the current posterior distribution.
+        \State Observe $L_{\textrm{best}} = \mathcal{L}(\mathcal{D}|\lambda_{\textrm{best}})$.
         \State Increment $n$
